---
title: "Paper charts"
output:
  html_document:
    theme: paper
    toc: true
    toc_float:
      collapsed: false
---

<style>
.main-container {
  max-width: 95% !important;
}
.toc-content {
  padding-left: 0px !important;
}
.svg-container {
  margin: auto !important;
}
</style>


<!-- Setting default chunk options -->
```{r chunkOptions, echo=FALSE, include = FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# setting global R chunk options (https://yihui.name/knitr/options/#chunk_options)
  knitr::opts_chunk$set(dev='svglite', # svg, png,...
                        fig.ext = ".svg",
                        #fig.asp = .8 # default aspect ratio
                        fig.width = 12,
                        fig.height = 6,
                        dpi=100
                        )
```  


```{r loading_packages, echo=FALSE, include = FALSE}

  #Loading required packages
  packagesList <- c("reticulate", "tidyr", "dplyr", "quitte", "openxlsx", "piamInterfaces", "ggplot2", "grid", "gridExtra", "ggh4x", "ggrepel")
  packages <- suppressWarnings(suppressMessages(lapply(packagesList, function(x){ if(!require(x, character.only = T, quietly = T)){ install.packages(x); return(paste0("installed missing package: ", x)) } else paste0("required package is installed: ", x) } )))

  #use_python("C:/Users/robertp/AppData/Local/Programs/Python/Python313")
```


<!-- ## download and prepare data -->

```{r load_data, echo=FALSE, include = FALSE}

downloadData <- FALSE
filterChartsData <- FALSE   # TRUE narrows df to the Figure-1..6 subset - do not use for a full run
rePrepareData <- FALSE

# define models groups
modelsFullSystem   <- c("IMAGE","PRIMES","PROMETHEUS","REMIND","TIAM-ECN","WITCH")
modelsElecall      <- c("IMAGE","LIMES", "MEESA","OSeMBE","PRIMES","PROMETHEUS","REMIND","TIAM-ECN","WITCH","Euro-Calliope")

# select data file
dataFolder <- sort(list.dirs(path = "./data", full.names = TRUE, recursive = FALSE), decreasing = T)[1]
time <- basename(dataFolder)
outFolder <- paste0("./output/", time)
dataRDSFile <- paste0("./data/", time, "/WP1_", time, ".rds")
if(!(downloadData) & !(rePrepareData) &(file.exists(dataRDSFile))){
  if(filterChartsData){
    df <- readRDS("./data/results.rds")
    histData <- readRDS("./data/historical.rds")
  } else {
    df <- readRDS(dataRDSFile)
    histRDS <- paste0("./data/", time, "/WP1_", time, "_hist.rds")
    histData <- readRDS(histRDS)
  }
} else {
  source("./prepare_data.R")
  dataFolder <- sort(list.dirs(path = "./data", full.names = TRUE, recursive = FALSE), decreasing = T)[1]
  time <- basename(dataFolder)
  outFolder <- paste0("./output/", time)
  dataRDS <- paste0("./data/", time, "/WP1_", time, ".rds")
  saveRDS(df, dataRDS)
  histRDS <- paste0("./data/", time, "/WP1_", time, "_hist.rds")
  histData <- histOrigN
  saveRDS(histData, histRDS)
}

```

<!-- ## paper calculations -->

## GHG emission (synthetic)  reductions vs 1990 w/ intra-EU bunkers - calc
```{r ghg_emissions_reductions_vs_1990_wIntraEU, echo=FALSE, include=FALSE}
# paper calculations

## ----------------- with intra-EU bunkers

GHGemiWintraEUB <- df %>%
  filter(variable == "Emi|Kyoto Gases|synthetic w/ intraEU Bunkers",
         period %in% c(2010,2020,2030,2035,2040,2050), scenario == "WP1 NetZero", region == "EU27 & UK (*)", model %in% modelsFullSystem )

reductionsVs1990_GHGemiWintraEUB <- mutate(GHGemiWintraEUB, value = 1 - value/5453)

quartiles_reductionsVs1990_GHGemiWintraEUB <- reductionsVs1990_GHGemiWintraEUB %>%
  group_by(period) %>%
  summarise_at(vars(value), list(
    Min = min,
    Mean = mean,
    Median = median,
    Max = max,
    Sd = sd,
    Q1=~quantile(., probs = 0.25),
    Q3=~quantile(., probs = 0.75)))

## ----------------- with intra-EU aviation

GHGemiWintraEUAv <- df %>%
  filter(variable == "Emi|Kyoto Gases|synthetic w/ intraEU Aviation",
         period %in% c(2010,2020,2030,2035,2040,2050), scenario == "WP1 NetZero", region == "EU27 & UK (*)", model %in% modelsFullSystem )

reductionsVs1990_GHGemiWintraEUAv <- mutate(GHGemiWintraEUAv, value = 1 - value/5407)

quartiles_reductionsVs1990_GHGemiWintraEUAv <- reductionsVs1990_GHGemiWintraEUAv %>%
  group_by(period) %>%
  summarise_at(vars(value), list( Min = min, Mean = mean, Median = median, Max = max, Sd = sd, Q1=~quantile(., probs = 0.25), Q3=~quantile(., probs = 0.75)))

## ----------------- without bunkers

GHGemiwoBunk <- df %>%
  filter(variable == "Emi|Kyoto Gases|synthetic wo Bunkers",
         period %in% c(2010,2020,2030,2035,2040,2050), scenario == "WP1 NetZero", region == "EU27 & UK (*)", model %in% modelsFullSystem )

reductionsVs1990_GHGemiWoB <- df %>%
  filter(variable == "Emi|Kyoto Gases|synthetic wo Bunkers| reduction vs 1990",
         period %in% c(2010,2020,2030,2035,2040,2050), scenario == "WP1 NetZero", region == "EU27 & UK (*)", model %in% modelsFullSystem )
quartiles_reductionsVs1990_GHGemiWoB <- reductionsVs1990_GHGemiWoB %>%
  group_by(period) %>%
  summarise_at(vars(value), list(
    Min = min,
    Mean = mean,
    Median = median,
    Max = max,
    Sd = sd,
    Q1=~quantile(., probs = 0.25),
    Q3=~quantile(., probs = 0.75)))

## ----------------- with all bunkers

reductionsVs1990_GHGemiWallEUB <- df %>%
  filter(variable == "Emi|Kyoto Gases|synthetic| reduction vs 1990",
         period %in% c(2010,2020,2030,2035,2040,2050), scenario == "WP1 NetZero", region == "EU27 & UK (*)", model %in% modelsFullSystem )



quartiles_reductionsVs1990_GHGemiWallEUB <- reductionsVs1990_GHGemiWallEUB %>%
  group_by(period) %>%
  summarise_at(vars(value), list(
    Min = min,
    Mean = mean,
    Median = median,
    Max = max,
    Sd = sd,
    Q1=~quantile(., probs = 0.25),
    Q3=~quantile(., probs = 0.75)))


plotDatafull <- df %>%
  filter(variable == "Emi|Kyoto Gases|synthetic w/ intraEU Bunkers| reduction vs 1990",
         period %in% c(2010,2020,2030,2035,2040,2050),
         scenario == "WP1 NetZero",
         region == "EU27 & UK (*)", 
         model %in% modelsFullSystem ) 
plotData <- plotDatafull %>%
  mutate (value = - value) %>%
  group_by(period) %>%
  summarise_at(vars(value), list(
    Min = min,
    Mean = mean,
    Median = median,
    Max = max,
    Sd = sd,
    Q1=~quantile(., probs = 0.25),
    Q3=~quantile(., probs = 0.75)))
    
```


```{r create_charts, echo=FALSE, include=FALSE}

  # creating charts
  g <- NULL

  # output folder
  dir.create(paste0("./output/", time, "/svg"), recursive = TRUE, showWarnings = FALSE)
  dir.create(paste0("./output/", time, "/png"), recursive = TRUE, showWarnings = FALSE)

``` 


```{r aesthetics, echo=FALSE, include=FALSE}

color <- c(
  #model
  "Historical" = "black",
  "mean" = "#c0c0c0",
  "median" = "#555555",
  "IMAGE" = "#00ffff",
  "Euro-Calliope" = "#84eab3",
  "MESSAGE" = "#800080",
  "PRIMES" = "#0000ff",
  "PROMETHEUS" = "#ffd900",
  "REMIND" = "#ff6347",
  "TIAM-ECN" = "#4682b4",
  "TIAM-ECN 1.2" = "#000000",
  "WITCH" = "#228b22",
  "LIMES" = "#ffaf47",
  "MEESA" = "#ff00ff",
  "OSeMBE" = "#a52a2a"
)

modelLinetype <- c(
  "Historical" = "solid",
  "mean" = "42", # dashed, 4 points on, one off
  "median" = "solid",
  "IMAGE" =       "solid" ,
  "Euro-Calliope" = "solid", 
  "MESSAGE" =    "solid",
  "PRIMES" =     "solid",
  "PROMETHEUS" = "solid",
  "REMIND" =     "solid",
  "TIAM-ECN" =   "solid",
  "TIAM-ECN 1.2" = "solid",
  "WITCH" =      "solid",
  "LIMES" =      "solid",
  "MEESA" =      "solid",
  "OSeMBE" =     "solid"
)

modelAlpha <- c(
  "Historical" =      1,
  "mean" =            1,
  "median" =          1,
  "IMAGE" =         0.5,
  "Euro-Calliope" = 0.5,
  "MESSAGE" =       0.5,
  "PRIMES" =        0.5,
  "PROMETHEUS" =    0.5,
  "REMIND" =        0.5,
  "TIAM-ECN" =      0.5,
  "TIAM-ECN 1.2" =  0.5,
  "WITCH" =         0.5, 
  "LIMES" =         0.5,
  "MEESA" =         0.5,
  "OSeMBE" =        0.5
)

modelsOrder <- c(
  "Historical",
  "mean",
  "median",
  "OSeMBE",
  "MEESA",
  "LIMES",
  "WITCH",
  "TIAM-ECN",
  "TIAM-ECN 1.2",
  "REMIND",
  "PROMETHEUS",
  "PRIMES",
  "MESSAGE",
  "IMAGE",
  "Euro-Calliope"
)

modelShape <- c(
  "Historical"    = NA, 
  "mean"          = NA, 
  "median"        = 21,
  "IMAGE"         = 21,
  "Euro-Calliope" = 21, 
  "MESSAGE"       = 21, 
  "PRIMES"        = 21, 
  "PROMETHEUS"    = 21, 
  "REMIND"        = 21, 
  "TIAM-ECN"      = 21, 
  "TIAM-ECN 1.2"  = 21, 
  "WITCH"         = 21,
  "LIMES"         = 23, 
  "MEESA"         = 23, 
  "OSeMBE"        = 23 
)

``` 


```{r paper_theme, echo=FALSE, include=FALSE}

theme_paper <- function(){
  theme_minimal(
    base_size = 24
    #, base_family = "Helvetica"
    ) +
    theme(
      # Panel background and grid
      panel.background = element_rect(fill = "white", colour = NA),
      panel.grid.major = element_line(colour = "grey90", size = .7),
      panel.grid.minor = element_line(colour = "grey95", size = .7),
      plot.background = element_rect(fill="#FFFFFF", color = NA),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.margin = margin(r = 20, l = 20),
      # Axis lines, titles, and text
      axis.ticks = element_line(colour = "black", size = 0.5),
      axis.title = element_text(size = 14, face = "plain", colour = "black"),
      axis.text = element_text(size = 12, colour = "black"),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),	  
      # Plot title and subtitle
      plot.title = element_text(size = 16, face = "plain", hjust = 0, margin = margin(t = 10, b = 10), colour = "black"),
      plot.subtitle = element_text(size = 12, face = "plain", hjust = 0, margin = margin(b = 10), colour = "grey50"),	  
      # Legend
      legend.position = "bottom",
      legend.box = "horizontal",
      legend.title = element_blank(),
      legend.text = element_text(size = 12, colour = "black"),
      #legend.key.size = unit(0.5, "lines"),
      legend.key.width = unit(2, 'cm'),  
      # Strip for facets
      strip.background = element_rect(fill = "white", colour = NA),
      #strip.text = element_blank(),
      panel.spacing = unit(3, "lines")
    )
}

def <- list(
  grid = list(
    lineSize = .7,
    color = "grey90",
    linetype = "dashed"
  ),
  yAxis= list(
    linewidth = 2,
    alpha = 0.3
  ),
  hist = list(
    lineSize = 1.5,
    areaColor = "#eeeeee",
    alpha = 0.5
  ),
  model = list(
    lineSize = 0.8,
    pointSize = 0.5,
    pointStroke = 1.5,
    shape = 21,
    smallPointSize = 0.5,
    smallPointStroke = 1
  ),
  median = list(
    lineSize = 1.5,
    pointSize = 2,
    pointStroke = 2,
    shape = 21,
    smallPointSize = 1,
    smallPointStroke = 1
  ),
  decadeChange = list(
    lineSize = 0.5,
    wideLineSize = 0.5,
    wideAlpha = 0.3,
    pointSize = 1,
    pointStroke = 1
  )
)

# End of the grey "Historical" band. Follow the data instead of a hardcoded year, so the
# band can never run past the last historical observation - the sources end in different
# years (UNFCCC 2022, Ember 2024, IEA 2023) and they move with each historical.mif.
# `fallback` is used for panels that carry no historical series of their own.
histEnd <- function(d, fallback = 2024) {
  p <- d$period[d$model == "Historical"]
  p <- p[!is.na(p)]
  if (!length(p)) fallback else max(p)
}

dir.create(paste0(outFolder, "/png"), showWarnings = FALSE, recursive = TRUE)
dir.create(paste0(outFolder, "/svg"), showWarnings = FALSE, recursive = TRUE)

```


## Figure 1

<!-- ### Figure 1.a -->

<!-- Greenhouse gas emissions for scenarios reaching climate neutrality by 2050 -->

<!-- Decadal average of the annual percentage change in the median model results, expressed relative to 1990 baseline values, shown in blue. -->

```{r figure_1, echo=FALSE, include=FALSE}

plotDataRaw <- df %>%
  filter(variable == "Emi|Kyoto Gases|synthetic",
         period >=2010, period <= 2050,
         scenario %in% c("WP1 NetZero"),
         model %in% modelsFullSystem, 
         region == "EU27 & UK (*)") %>%
  rbind(
    df %>%
      filter(variable == "Emissions|CO2|Energy and Industrial Processes",
             period >=2010, period <= 2050,
             model %in% modelsFullSystem,
             scenario %in% c("WP1 NetZero"),
             region == "EU27 & UK (*)")
  ) %>%
  rbind(
    histData %>%
      filter(variable == "Emi|GHG|w/ Bunkers|LULUCF national accounting",
             region == "EUR",
             model == "UNFCCC") %>%
      mutate(variable = "Emi|Kyoto Gases|synthetic",
             model = "Historical") %>%
      rbind(
        histData %>%
          filter(variable == "Emi|CO2|w/ Bunkers|Energy and Industrial Processes",
                 region == "EUR",
                 model == "UNFCCC") %>%
          mutate(variable = "Emissions|CO2|Energy and Industrial Processes",
                 model = "Historical")
      )
  )

plotData <- plotDataRaw %>%
  rbind(
    plotDataRaw %>%
      filter(model != "Historical") %>%
      group_by(scenario, region, variable, unit, period) %>%
      summarise(median_value = median(value, na.rm = TRUE)) %>%
      ungroup() %>%
      rename(value = median_value) %>%
      mutate(model = "median") %>%
      select(all_of(names(plotDataRaw)))
  ) %>%
  mutate(model = factor(model, levels=rev(modelsOrder)))

decade_change <- plotData %>%
  filter(model == "median", period %% 10 == 0, period > 2020) %>%
  rbind(
    plotData %>%
      filter(model == "Historical", period %% 10 == 0, period <= 2020)
  ) %>%
  select(period, variable, value) %>%
  arrange(variable,period) %>%
  group_by(variable) %>%
  mutate(
    change_in_10_years = value - lag(value),
    change_percentage = ((change_in_10_years) / value[period == 1990])/10 * 100,
    prev_period = lag(period),
    prev_value = lag(value)
  ) %>%
  #filter(!is.na(prev_period)) %>%
  mutate(
    x = (period + prev_period) / 2,
    y = (value + prev_value) / 2,
    vAlign = case_when(
      period == 2000 ~ -0.5, 
      period == 2020 ~ -2.5, 
      period %in% c(2030, 2040) ~ -3.0,
      #period == 2050 ~ 1.5,
      TRUE ~ -1.5
    )
  )

limitY <- max(plotData$value)

g_data <- list(
  "Emi|Kyoto Gases|synthetic" = list(
    data = plotData %>% filter(variable == "Emi|Kyoto Gases|synthetic"),
    decade_change = decade_change %>% filter(variable == "Emi|Kyoto Gases|synthetic"),
    title = "GHG Emissions",
    subtitle = expression(paste("Mt ", CO[2]," eq/yr"))
  ),
  "Emissions|CO2|Energy and Industrial Processes" = list(
    data = plotData %>% filter(variable == "Emissions|CO2|Energy and Industrial Processes"),
    decade_change = decade_change %>% filter(variable == "Emissions|CO2|Energy and Industrial Processes"),
    title = expression(paste(CO[2],"  Emissions - energy and industrial processes")),
    subtitle = expression(paste("Mt ", CO[2]," /yr"))
  )
)

g <- list()
for( chart in names(g_data)){
  g[[chart]] <- ggplot(g_data[[chart]][["data"]],aes(x=period,y=value)) +
    geom_segment(data= g_data[[chart]][["data"]] %>% filter(period %% 10 == 0), aes(x = period, xend = period, y = 0, yend = value), color = def$grid$color, size = def$grid$lineSize, linetype = def$grid$linetype) + # vertical grid line
    annotate("rect", xmin = -Inf, xmax = histEnd(g_data[[chart]][["data"]]), ymin = -Inf, ymax = Inf, fill = def$hist$areaColor, alpha = def$hist$alpha) + # historical area
    annotate("text", x = 2005, y = 3000, label = "Historical", hjust = 0.5, vjust = 0.5, size = 15/.pt, color = "#888888") + # historical text
    geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
    geom_line(data=g_data[[chart]][["data"]] %>% filter((!(model %in% c("Historical", "median")))), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
    geom_point(data = g_data[[chart]][["data"]] %>% filter(!(model %in% c("Historical", "median")), period > 2023), shape = def$median$shape, aes(colour = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) +
    geom_line(data=g_data[[chart]][["data"]] %>% filter(model == "median"), size=def$median$lineSize,aes(color=model,linetype=model, alpha = model)) +
    geom_point(data = g_data[[chart]][["data"]] %>% filter(model == "median"), shape = def$median$shape, aes(colour = model), fill = "white", size = def$median$pointSize, stroke = def$median$pointStroke) +
    geom_line(data=g_data[[chart]][["data"]] %>% filter(model == "Historical"), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) +
    geom_label(data = g_data[[chart]][["decade_change"]] %>% filter(!is.na(prev_value)), aes(x = x, y = y, label = paste0(round(change_percentage, 1), "% / yr"), vjust = vAlign), fill = "#90D5FF", color = "black", label.padding = unit(0.25, "lines"), label.r = unit(0.15, "lines"), hjust = 0.5, label.size = 0, alpha = 0.5, size = 2) +
    theme_paper() +
    facet_wrap(~ variable) + 
    scale_color_manual(name = "legend", values = setNames(color[levels(plotData$model)],levels(plotData$model)), guide = guide_legend(nrow = 2)) +
    scale_alpha_manual(values = modelAlpha) +
    scale_linetype_manual(values = modelLinetype) +
    labs(title = g_data[[chart]][["title"]], subtitle = g_data[[chart]][["subtitle"]]) + 
    scale_x_continuous(expand = c(0, 0)) +
    expand_limits(y = limitY) +
    coord_cartesian(clip = 'off') +
    theme(strip.text = element_blank()) +
    guides(linetype = "none", alpha = "none")
}

# function to extract the legend
g_legend <- function(a.gplot){ 
    tmp <- ggplot_gtable(ggplot_build(a.gplot)) 
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box") 
    legend <- tmp$grobs[[leg]] 
    legend
} 

g[["legend"]] <- g_legend(g$`Emi|Kyoto Gases|synthetic` + theme(plot.margin = margin(t = 20)))

g[["Figure 1.a"]] <- gridExtra::arrangeGrob(
  g$`Emi|Kyoto Gases|synthetic` + theme(legend.position="none"),
  g$`Emissions|CO2|Energy and Industrial Processes` + theme(legend.position="none"),
  g[["legend"]],
  layout_matrix = rbind(c(1, 2),
                        c(1, 2),
                        c(1, 2),
                        c(1, 2),
                        c(1, 2),
                        c(3, 3))
   )

```

<!-- ### Figure 1.b -->

<!-- Electricity Supply CO2 emissions -->

<!-- Decadal average of the annual percentage change in the median model results, expressed relative to 2000 baseline values, shown in blue. -->

```{r figure 1.b, echo=FALSE, include=FALSE}

plotDataRaw <- df %>%
  filter(variable == "Emissions|CO2|Energy|Supply|Electricity",
         period >=2010, period <= 2050,
         scenario %in% c("WP1 NetZero"),
         model %in% modelsElecall, 
         region == "EU27 & UK (*)") %>%
  rbind(
    histData %>%
    filter(variable == "Emi|CO2|Energy|Supply|Electricity w/ couple prod",
           region == "EUR",
           model == "Ember") %>%
      mutate(variable = "Emissions|CO2|Energy|Supply|Electricity",
             model = "Historical")
    )

plotData <- plotDataRaw %>%
  rbind(
    plotDataRaw %>%
      filter(!model %in% c("Historical", "Euro-Calliope")) %>% # Euro-Calliope excluded from the median
      group_by(scenario, region, variable, unit, period) %>%
      summarise(median_value = median(value, na.rm = TRUE)) %>%
      ungroup() %>%
      rename(value = median_value) %>%
      mutate(model = "median") %>%
      select(all_of(names(plotDataRaw)))
  ) %>%
  mutate(model = factor(model, levels=rev(modelsOrder)))

decade_change <- plotData %>% 
  filter(model == "median", period %% 10 == 0, period > 2020) %>%
  rbind(
    plotData %>%
      filter(model == "Historical", period %% 10 == 0, period <= 2020)
  ) %>%
  select(period, variable, value) %>%
  arrange(variable,period) %>%
  group_by(variable) %>%
  mutate(
    change_in_10_years = value - lag(value),
    change_percentage = ((change_in_10_years) / value[period == 2000])/10 * 100,
    prev_period = lag(period),
    prev_value = lag(value)
  ) %>%
  mutate(
    x = (period + prev_period) / 2,
    y = (value + prev_value) / 2,
    vAlign = case_when(
      #period == 2040 ~ 1.8,
      period %in% c(2020,2030) ~ -3.0,
      TRUE ~ -1.5
    )
  )

g[["Figure 1.b"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_segment(data = plotData %>% filter(period %% 10 == 0), aes(x = period, xend = period, y = 0, yend = value), color = def$grid$color, size = def$grid$lineSize, linetype = def$grid$linetype) + # vertical grid line
  annotate("rect", xmin = -Inf, xmax = histEnd(plotData), ymin = -Inf, ymax = Inf, fill = def$hist$areaColor, alpha = def$hist$alpha) + # historical area
  annotate("text", x = 2010, y = 750, label = "Historical", hjust = 0.5, vjust = 0.5, size = 15/.pt, color = "#888888") + # historical text
  geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
  geom_line(data=plotData %>% filter((!(model %in% c("Historical", "median", "Euro-Calliope")))), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(!model %in% c("median", "Euro-Calliope"), period > 2023), aes(colour = model, shape = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) +
  geom_point(data = plotData %>% filter(model == "Euro-Calliope", period %in% c(2020, 2030, 2050)), aes(colour = model, shape = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) + # Euro-Calliope as symbols only, no line
  geom_line(data=plotData %>% filter(model == "median"), size=def$median$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(model == "median"), aes(colour = model, shape = model), fill = "white", size = def$median$pointSize, stroke = def$median$pointStroke) +
  geom_line(data=plotData %>% filter(model == "Historical"), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + 
  geom_label(data = decade_change %>% filter(!is.na(prev_value)), aes(x = x, y = y, label = paste0(round(change_percentage, 1), "% / yr"), vjust = vAlign), fill = "#90D5FF", color = "black", label.padding = unit(0.25, "lines"), label.r = unit(0.15, "lines"), hjust = 0.5, label.size = 0, alpha = 0.5, size = 2) +
  theme_paper() + 
  scale_color_manual(name = "legend", values = setNames(color[levels(plotData$model)], levels(plotData$model)), guide = guide_legend(ncol = 2)) +
  scale_alpha_manual(values = modelAlpha) +
  scale_linetype_manual(values = modelLinetype) +
  scale_shape_manual(name = "legend", values = modelShape, guide = guide_legend(ncol = 2)) +
  labs(title = expression(paste("Electricity Supply ", CO[2]," Emissions")), subtitle = expression(paste("Mt ", CO[2]," /yr"))) +
  scale_x_continuous(expand = c(0, 0)) +
  coord_cartesian(clip = 'off') + 
  theme(legend.position = "right",
        legend.key.height = unit(1, 'cm'),
        plot.margin = margin(r = 100, l = 100),
        legend.margin = margin(l = 50),
        strip.text = element_blank()
  ) +
  guides(linetype = "none", alpha = "none")

```

<!-- ### Figure 1 -->

```{r figure 1, echo=FALSE, include=FALSE}

g[["Figure 1"]] <- gridExtra::arrangeGrob(
  g$`Emi|Kyoto Gases|synthetic` + 
    theme(legend.position="none") + 
    labs(title = "(a) Net GHG emissions", subtitle = expression(paste("Mt ", CO[2]," eq/yr"))) +
    theme(plot.margin = margin(b = 10, r=20)),
  g$`Emissions|CO2|Energy and Industrial Processes` +
    theme(legend.position="none") +
    labs(title = expression(paste("(b) Net ", CO[2], " emissions - energy and industrial processes")), subtitle = expression(paste("Mt ", CO[2]," eq/yr"))) +
    theme(plot.margin = margin(b = 10, r=20),
          plot.title = element_text(hjust = 0.2)) +
      coord_cartesian(clip = "off"),
  g[["Figure 1.b"]] + 
    theme(plot.margin = margin(r = 60, t = 10), legend.margin = margin(l = 80)) + 
    labs(title = expression(paste("(c) Net ", CO[2]," emissions from electricity generation")), subtitle = expression(paste("Mt ", CO[2]," eq/yr"))),
  layout_matrix = rbind(c(1, 2),
                        c(3, 3))
   )

ggsave(paste0(outFolder, "/png/Figure 1 Emissions panel.png"), g[["Figure 1"]], device="png", width = 12, height = 10, dpi=300, units = "in")
ggsave(paste0(outFolder, "/svg/Figure 1 Emissions panel.svg"), g[["Figure 1"]], device="svg", width = 12, height = 10, dpi=100, units = "in")

```


```{r, echo=FALSE, fig.width=12, fig.height=10}

grid.newpage()
grid.draw(g[["Figure 1"]])

```

Figure 1. Emissions panel (EU27 and UK). Historical UNFCCC (figure 1.a and 1.b) and Ember (Figure 1.c) data shown in black. Decadal average of the annual percentage change in the median model results, expressed relative to 1990 (figure 1.a and 1.b) or 2000 (figure 1.c) baseline values, shown in blue.

---

## Figure 2

<!-- Electricity Generation from Wind and Solar and Share in electricity generation -->

```{r SE_Electricity_VRE, echo=FALSE, include=FALSE}

plotDataRaw <- df %>%
  filter(variable == "Secondary Energy|Electricity|WindSolar",
         period >=2010, period <= 2050,
         scenario %in% c("WP1 NetZero"),
         model %in% modelsElecall, 
         region == "EU27 & UK (*)") %>%
  rbind(
    histData %>%
    filter(variable == "SE|Electricity|WindSolar",
           region == "EUR",
           model == "Ember") %>%
      mutate(variable = "Secondary Energy|Electricity|WindSolar",
             model = "Historical")
    ) %>%
  mutate(value = value*277.7777778) # EJ 2 TWh

plotData <- plotDataRaw %>%
  rbind(
    plotDataRaw %>%
      filter(!model %in% c("Historical", "Euro-Calliope")) %>% # Euro-Calliope excluded from the median
      group_by(scenario, region, variable, unit, period) %>%
      summarise(median_value = median(value, na.rm = TRUE)) %>%
      ungroup() %>%
      rename(value = median_value) %>%
      mutate(model = "median") %>%
      select(all_of(names(plotDataRaw)))
  ) %>%
  mutate(model = factor(model, levels=rev(modelsOrder)))

decade_change <- plotData %>% 
  filter(model == "median", period %% 10 == 0, period > 2020) %>%
  rbind(
    plotData %>%
      filter(model == "Historical", period %% 10 == 0, period <= 2020)
  ) %>%
  select(period, variable, value) %>%
  arrange(variable,period) %>%
  group_by(variable) %>%
  mutate(
    change_in_10_years = value - lag(value),
    change_percentage = ((change_in_10_years) / lag(value))/10*100,
    prev_period = lag(period),
    prev_value = lag(value)
  ) %>%
  #filter(!is.na(prev_period)) %>%
  mutate(
    x = (period + prev_period) / 2,
    y = (value + prev_value) / 2,
    vAlign = case_when(
      TRUE ~ -1.5
    )
  )

g[["Electricity Generation from Wind and Solar"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_segment(data = plotData %>% filter(period %% 10 == 0), aes(x = period, xend = period, y = 0, yend = value), color = def$grid$color, size = def$grid$lineSize, linetype = def$grid$linetype) + # vertical grid line
  annotate("rect", xmin = -Inf, xmax = histEnd(plotData), ymin = -Inf, ymax = Inf, fill = def$hist$areaColor, alpha = def$hist$alpha) + # historical area
  annotate("text", x = 2012, y = 5000, label = "Historical", hjust = 0.5, vjust = 0.5, size = 15/.pt, color = "#888888") + # historical text
  geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
  geom_line(data = plotData %>% filter((!(model %in% c("Historical", "median", "Euro-Calliope")))),size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(period > 2023, !model %in% c("median", "Euro-Calliope")), aes(colour = model, shape = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) +
  geom_point(data = plotData %>% filter(model == "Euro-Calliope", period %in% c(2020, 2030, 2050)), aes(colour = model, shape = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) + # Euro-Calliope as symbols only, no line
  geom_line(data=plotData %>% filter(model == "median"), size=def$median$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(model == "median"), aes(colour = model, shape = model), fill = "white", size = def$median$pointSize, stroke = def$median$pointStroke) +
  geom_line(data=plotData %>% filter(model == "Historical"), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + 
  theme_paper() +
  scale_color_manual(name = "legend",values = setNames(color[levels(plotData$model)],levels(plotData$model)), guide = guide_legend(nrow = 2)) +
  scale_alpha_manual(values = modelAlpha) +
  scale_linetype_manual(values = modelLinetype) +
  scale_shape_manual(name = "legend",values = modelShape, guide = guide_legend(nrow = 2)) +
  labs(title = "(a) Electricity generation from wind and solar", subtitle = "TWh/yr") + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 8000, by = 2000), limits = c(0,9000), expand = c(0, 0)) +
  coord_cartesian(clip = 'off') +
  theme(strip.text = element_blank()) +
  expand_limits(y = c(0,1)) +
  guides(linetype = "none", alpha = "none")

plotDataRaw <- df %>%
  filter(variable == "Secondary Energy|Electricity|VRE Share",
         period >=2010, period <= 2050,
         scenario %in% c("WP1 NetZero"),
         model %in% modelsElecall, 
         region == "EU27 & UK (*)") %>%
  rbind(
    histData %>%
    filter(variable == "SE|Electricity|VRE Share",
           region == "EUR",
           model == "Ember") %>%
      mutate(variable = "Secondary Energy|Electricity|VRE Share",
             model = "Historical")
    )


plotData <- plotDataRaw %>%
  rbind(
    plotDataRaw %>%
      filter(!model %in% c("Historical", "Euro-Calliope")) %>% # Euro-Calliope excluded from the median
      group_by(scenario, region, variable, unit, period) %>%
      summarise(median_value = median(value, na.rm = TRUE)) %>%
      ungroup() %>%
      rename(value = median_value) %>%
      mutate(model = "median") %>%
      select(all_of(names(plotDataRaw)))
  ) %>%
  mutate(model = factor(model, levels=rev(modelsOrder)))

decade_change <- plotData %>% 
  filter(model == "median", period %% 10 == 0, period > 2020) %>%
  rbind(
    plotData %>%
      filter(model == "Historical", period %% 10 == 0, period <= 2020)
  ) %>%
  select(period, variable, value) %>%
  arrange(variable,period) %>%
  group_by(variable) %>%
  mutate(
    change_in_10_years = value - lag(value),
    change_percentage = ((change_in_10_years) / lag(value))/10*100,
    prev_period = lag(period),
    prev_value = lag(value)
  ) %>%
  mutate(
    x = (period + prev_period) / 2,
    y = (value + prev_value) / 2,
    vAlign = case_when(
      TRUE ~ -1.5
    )
  )

g[["Share in electricity generation"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_segment(data = plotData %>% filter(period %% 10 == 0), aes(x = period, xend = period, y = 0, yend = value), color = def$grid$color, size = def$grid$lineSize, linetype = def$grid$linetype) + # vertical grid line
  annotate("rect", xmin = -Inf, xmax = histEnd(plotData), ymin = -Inf, ymax = Inf, fill = def$hist$areaColor, alpha = def$hist$alpha) + # historical area
  annotate("text", x = 2012, y = 0.5, label = "Historical", hjust = 0.5, vjust = 0.5, size = 15/.pt, color = "#888888") + # historical text
  geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
  geom_line(data=plotData %>% filter((!(model %in% c("Historical", "median", "Euro-Calliope")))), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(period > 2023, !model %in% c("median", "Euro-Calliope")), aes(colour = model, shape = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) +
  geom_point(data = plotData %>% filter(model == "Euro-Calliope", period %in% c(2020, 2030, 2050)), aes(colour = model, shape = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) + # Euro-Calliope as symbols only, no line
  geom_line(data=plotData %>% filter(model == "median"), size=def$median$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(model == "median"), aes(colour = model, shape = model), fill = "white", size = def$median$pointSize, stroke = def$median$pointStroke) +
  geom_line(data=plotData %>% filter(model == "Historical"), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + 
  theme_paper() +
  scale_color_manual(name = "legend", values = setNames(color[levels(plotData$model)],levels(plotData$model)), guide = guide_legend(nrow = 2)) +
  scale_alpha_manual(values = modelAlpha) +
  scale_linetype_manual(values = modelLinetype) +
  scale_shape_manual(name = "legend", values = modelShape, guide = guide_legend(nrow = 2)) +
  labs(title = "(b) Share of wind and solar in electricity generation", subtitle = "%") + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, 1, by = 0.2), limits = c(0,1), expand = c(0, 0)) +
  coord_cartesian(clip = 'off') +
  theme(strip.text = element_blank()) +
  expand_limits(y = c(0,1)) +
  guides(linetype = "none", alpha = "none")

g[["legend"]] <- g_legend(g$`Electricity Generation from Wind and Solar` + guides(color = guide_legend(nrow = 2)) + theme(plot.margin = margin(t = 20)))

g[["Figure 2"]] <- gridExtra::arrangeGrob(
  g$`Electricity Generation from Wind and Solar` + theme(legend.position="none", plot.margin = margin(r=10, l=20)),
  g$`Share in electricity generation` + theme(legend.position="none", plot.margin = margin(r=20, l=10, b=15), 
                                              #plot.subtitle = element_text(margin = margin(b = 40))
                                              ),
  g[["legend"]],
  layout_matrix = rbind(c(1, 2),
                        c(1, 2),
                        c(1, 2),
                        c(1, 2),
                        c(1, 2),
                        c(3, 3))
   )


ggsave(paste0(outFolder, "/png/Figure 2. Electricity Generation from Wind and Solar and Share in electricity generation.png"), g[["Figure 2"]], device="png", width = 12, height = 6, dpi=300, units = "in")
ggsave(paste0(outFolder, "/svg/Figure 2. Electricity Generation from Wind and Solar and Share in electricity generation.svg"), g[["Figure 2"]], device="svg", width = 12, height = 6, dpi=100, units = "in")


```


```{r, echo=FALSE, fig.width=12, fig.height=6}

grid.newpage()
grid.draw(g[["Figure 2"]])

```

Figure 2. Variable renewables generation and electricity share. Decadal average of the annual percentage change in the median model results, expressed relative to the preceding decade, shown in blue.


---


## Figure 3

<!-- Share of electricity and hydrogen in final energy -->

```{r fe_electricity_and_hydrogen_share, echo=FALSE, include=FALSE}

plotDataRaw <- df %>%
  filter(variable %in% c("Final Energy|Electricity Share woBunkers woNonEnergy", "Final Energy|Hydrogen Share woBunkers woNonEnergy"),
         period >=2010, period <= 2050,
         scenario %in% c("WP1 NetZero"),
         model %in% modelsFullSystem, 
         region == "EU27 & UK (*)") %>%
  rbind(
    histData %>%
    filter(variable == "Final Energy|Electricity Share woBunkers woNonEnergy",
           region == "EUR",
           period >=2000,
           model == "IEA") %>%
      mutate(variable = "Final Energy|Electricity Share woBunkers woNonEnergy",
             model = "Historical")
    )

plotData <- plotDataRaw %>%
  rbind(
    plotDataRaw %>%
      filter(model != "Historical") %>%
      group_by(scenario, region, variable, unit, period) %>%
      summarise(median_value = median(value, na.rm = TRUE)) %>%
      ungroup() %>%
      rename(value = median_value) %>%
      mutate(model = "median") %>%
      select(all_of(names(plotDataRaw)))
  ) %>%
  mutate(model = factor(model, levels=rev(modelsOrder)),
         carrier = ifelse(variable == "Final Energy|Electricity Share woBunkers woNonEnergy", "Electricity", "Hydrogen")
  )
  

decade_change <- plotData %>% 
  filter(model == "median", period %% 10 == 0, period > 2020) %>%
  rbind(
    plotData %>%
      filter(model == "Historical", period %% 10 == 0, period <= 2020)
  ) %>%
  select(period, variable, value) %>%
  arrange(variable,period) %>%
  group_by(variable) %>%
  mutate(
    change_in_10_years = value - lag(value),
    change_percentage = ((change_in_10_years) / lag(value))/10*100,
    prev_period = lag(period),
    prev_value = lag(value)
  ) %>%
  mutate(
    x = (period + prev_period) / 2,
    y = (value + prev_value) / 2,
    vAlign = case_when(
      period %in% c(2010, 2020) ~ - 0.5,
      TRUE ~ -1.5
    )
  )

 g[["Figure 3"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_segment(data = plotData %>% filter(period %% 10 == 0), aes(x = period, xend = period, y = 0, yend = value), color = def$grid$color, size = def$grid$lineSize, linetype = def$grid$linetype) + # vertical grid line
  annotate("rect", xmin = -Inf, xmax = histEnd(plotData), ymin = -Inf, ymax = Inf, fill = def$hist$areaColor, alpha = def$hist$alpha) + # historical area
  annotate("text", x = 2012, y = 0.5, label = "Historical", hjust = 0.5, vjust = 0.5, size = 15/.pt, color = "#888888") + # historical text
  geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
  geom_line(data=plotData %>% filter((!(model %in% c("Historical", "median")))), size=def$model$lineSize,aes(color=model,linetype=carrier, alpha = model), show.legend = c(linetype = FALSE, colour = TRUE)) + # models data
  geom_point(data = plotData %>% filter(!(model %in% c("Historical", "median")), period > 2023), shape = def$model$shape, aes(colour = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) +
  geom_line(data=plotData %>% filter(model == "median"), size=def$median$lineSize,aes(color=model,linetype=carrier, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(model == "median"), shape = def$median$shape, aes(colour = model, shape = model), fill = "white", size = def$median$pointSize, stroke = def$median$pointStroke) +
  geom_line(data=plotData %>% filter(model == "Historical"), size=def$model$lineSize,aes(color=model, alpha = model), linetype = "solid") + 
  theme_paper() +
  scale_color_manual(name = "legend", values = setNames(color[levels(plotData$model)],levels(plotData$model))) +
  scale_alpha_manual(values = modelAlpha) +
  scale_linetype_manual(name = "legend2", values = c("Electricity" = "solid", "Hydrogen" = "dashed")) +
  labs(title = "Share of electricity and hydrogen in final energy consumption", subtitle = "%") +
  scale_x_continuous(expand = c(0, 0)) +
  coord_cartesian(clip = 'off') +
  theme(strip.text = element_blank()) +
  expand_limits(y = c(0,1)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, 0.7, by = 0.1), limits = c(0,0.75)) + 
  theme(plot.margin = margin(r = 125, l = 125)) +
  guides(
    color = guide_legend(order = 1, nrow = 2),
    linetype = guide_legend(order = 2, nrow = 2),
    alpha = "none")

 ggsave(paste0(outFolder, "/png/Figure 3. Share of electricity and hydrogen in final energy.png"), g[["Figure 3"]], device="png", width = 12, height = 6, dpi=300, units = "in")
 ggsave(paste0(outFolder, "/svg/Figure 3. Share of electricity and hydrogen in final energy.svg"), g[["Figure 3"]], device="svg", width = 12, height = 6, dpi=100, units = "in")
 
```



```{r, echo=FALSE, fig.width=12, fig.height=6}

g[["Figure 3"]]

```


Figure 3. Electricity and hydrogen final energy shares. 
<!-- Decadal average of the annual percentage change in the median model results, expressed relative to the preceding decade, shown in blue. -->


---


## Figure 4

<!-- Left: Carbon-containing final energy. Right: Energy imports  -->


```{r Carbon-containing final energy, echo=FALSE, include=FALSE}

  vars <- list(
  "Synfuel - Gases"   = "Final Energy|Gases|Electricity",
  "Synfuel - Liquids" = "Final Energy|Liquids|Electricity",
  "Biomass - Gases"   = "Final Energy|Gases|Biomass",
  "Biomass - Liquids" = "Final Energy|Liquids|Biomass",
  "Biomass - Solids"  = "Final Energy|Solids|Biomass",
  "Fossils - Gases"   = "Final Energy|Gases|Fossil",   
  "Fossils - Liquids" = "Final Energy|Liquids|Fossil",
  "Fossils - Solids"  = "Final Energy|Solids|Fossil")

  carrier_color <- c(  
    "Synfuel - Gases"   = "#5ed5f0",
    "Synfuel - Liquids" = "#0177d4",
    "Biomass - Gases"   = "#33ed27",
    "Biomass - Liquids" = "#23ae1a",
    "Biomass - Solids"  = "#005900",
    "Fossils - Gases"   = "#c8c8c8",
    "Fossils - Liquids" = "#727272",
    "Fossils - Solids"  = "#0c0c0c")
  
  varToCarrier <- setNames(gsub("\\d+$", "", names(unlist(vars))),unlist(vars))

  plotData <- df %>%
    filter(variable %in% names(varToCarrier),
           period %in% c(2020,2030,2040,2050),
           model %in% modelsFullSystem,
           scenario == "WP1 NetZero",
           region == "EU27 & UK (*)") %>%
    mutate(model = factor(model, levels=rev(modelsOrder)),
           carrier = factor(varToCarrier[as.character(variable)], levels = names(vars))) %>%
    group_by(model,scenario,region,carrier,period) %>%
    summarise(value = sum(value))
     
  minfe <- plotData %>% group_by(model,period) %>% summarize(value= sum(value)) %>% pull(value) %>% min()
  maxfe <- plotData %>% group_by(model,period) %>% summarize(value= sum(value)) %>% pull(value) %>% max()
  
  plotDataTotal <- df %>%
    filter(variable == "FE|SolLiqGas",
           model %in% modelsFullSystem,
           period %in% c(2020,2030,2040,2050),
           scenario == "WP1 NetZero",
           region == "EU27 & UK (*)") %>%
    mutate(model = factor(model, levels=rev(modelsOrder)))
    
  g[["Carbon-containing final energy"]] <- ggplot(plotData) +
    #geom_rect(data = data.frame(period = 2020), xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "#eeeeee", alpha = 0.5) +
    geom_bar(stat='identity', alpha = 0.8, aes(x=model, y=value, fill=carrier)) +
    geom_point(data = plotDataTotal, aes(x=model, y = value), colour = "black", size = 3, shape = 18) +
    facet_wrap(~period,nrow=1) +
    theme_paper() +
    scale_fill_manual(values = carrier_color, guide = guide_legend(nrow = 3)) +
    theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5)) +
    labs(title = "(a) Carbon-based final energy", subtitle = "EJ/yr") +
    theme(legend.key.width = unit(0.7, 'cm'))
    
```

```{r Energy imports, echo=FALSE, include=FALSE}

  vars <- list(
    "Electricity" = "Trade|Secondary Energy|Electricity|Volume",
    "Hydrogen"    = "Trade|Secondary Energy|Hydrogen|Volume",
    "Liquids (Synfuel)" = c("Trade|Secondary Energy|Liquids|Electricity|Volume","Trade|Secondary Energy|Liquids|Hydrogen|Volume"),
    "Liquids (bio-based)" = "Trade|Secondary Energy|Liquids|Biomass|Volume",
    "Biomass" = c("Trade|Secondary Energy|Solids|Biomass|Volume", "Trade|Primary Energy|Biomass|Volume"),
    "Gas" = "Trade|Primary Energy|Gas|Volume",
    "Oil" = c("Trade|Secondary Energy|Liquids|Oil|Volume","Trade|Primary Energy|Oil|Volume"),
    "Coal - Liquids" = "Trade|Secondary Energy|Liquids|Coal|Volume",
    "Coal" = "Trade|Primary Energy|Coal|Volume"
  )

 carrier_color <- c(  
    "Electricity"         = "#fff300",
    "Hydrogen"            = "#00ffe4",
    "Liquids (Synfuel)"   = "#0177d4",
    "Liquids (bio-based)" = "#23ae1a",
    "Biomass"             = "#005900",
    "Gas"                 = "#b3b34b",
    "Oil"                 = "#f4786d",
    "Coal - Liquids"      = "#727272",
    "Coal"                = "#0c0c0c"
  ) 
 
  varToCarrier <- setNames(gsub("\\d+$", "", names(unlist(vars))),unlist(vars))

  histVars <- list(
    "Gas"  = "Trade|Gas",
    "Oil"  = "Trade|Oil",
    "Coal" = "Trade|Coal"
  )
  
  histVarToCarrier <- setNames(gsub("\\d+$", "", names(unlist(histVars))),unlist(histVars))
   
  plotData <- df %>%
    filter(variable %in% names(varToCarrier),
           period %in% c(2020,2030,2040,2050),
           model %in% modelsFullSystem,
           scenario == "WP1 NetZero",
           region == "EU27 & UK (*)") %>%
    mutate(carrier = factor(varToCarrier[as.character(variable)], levels = names(vars))) %>%
    rbind(
      histData  %>%
        filter(variable %in% names(histVarToCarrier),
               region == "EUR",
               period %in% c(2020,2030,2040,2050),
               model == "IEA") %>%
          mutate(carrier = factor(histVarToCarrier[as.character(variable)], levels = names(vars)),
                 model = "Historical")
      ) %>%
    group_by(model,scenario,region,carrier,period) %>%
    summarise(value = sum(value)) %>%
    mutate(model = factor(model, levels=rev(c(modelsOrder[modelsOrder != "Historical"], "Historical"))),
           value = value * - 1)
  
  minMport <- plotData %>% group_by(model,period) %>% summarize(value= sum(value)) %>% pull(value) %>% min()
  maxMport <- plotData %>% group_by(model,period) %>% summarize(value= sum(value)) %>% pull(value) %>% max()
  
  limitY <- c(min(minfe, minMport), max(maxfe,maxMport))
  
  g[["Energy imports"]] <- ggplot(plotData) +
    geom_bar(stat='identity', alpha = 0.8,aes(x=model,y=value,fill=carrier)) +
    facet_wrap(~period,nrow=1, scales = "free_x") +
    theme_paper() +
    scale_fill_manual(values = carrier_color, guide = guide_legend(nrow = 3)) +
    theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5)) +
    labs(title = "(b) Net energy imports", subtitle = "EJ/yr") +
    theme(legend.key.width = unit(0.7, 'cm')) +
    force_panelsizes(cols = plotData %>% select(model, period) %>% group_by(period) %>% summarize(unique_count = n_distinct(model)) %>% pull(unique_count))
	
    
```


```{r Carbon-containing final energy and energy imports, echo=FALSE, include=FALSE}

#g[["legend"]] <- g_legend(g$`Energy imports` + guides(color = guide_legend(nrow = 1)) + theme(plot.margin = margin(t = 20)))

g[["Figure 4"]] <- gridExtra::arrangeGrob(
  g$`Carbon-containing final energy` + 
    expand_limits(y = limitY),
  g$`Energy imports` +
    expand_limits(y = limitY),
  layout_matrix = rbind(c(1, 2))
   )

ggsave(paste0(outFolder, "/png/Figure 4. Carbon-containing final energy and energy imports.png"), g[["Figure 4"]], device="png", width = 12, height = 8, dpi=300, units = "in")
ggsave(paste0(outFolder, "/svg/Figure 4. Carbon-containing final energy and energy imports.svg"), g[["Figure 4"]], device="svg", width = 12, height = 8, dpi=100, units = "in")


```


```{r, echo=FALSE, fig.width=12, fig.height=8}

grid.newpage()
grid.draw(g[["Figure 4"]])

```


Figure 4. Residual carbon and energy imports.


---


## Figure 5

<!-- Emissions per sector -->


```{r Emissions per sector, echo=FALSE, include=FALSE}

  vars <- c(
    "Demand - Bunkers"            = "Emissions|CO2|Energy|Demand|Bunkers",
    "non-CO2"                     = "Emissions|Kyoto Gases|non-CO2 synthetic",
    "other CO2"                   = "Emissions|CO2|Other",
    "Demand - AFOFI"              = "Emissions|CO2|Energy|Demand|AFOFI",
    "Demand - Buildings"          = "Emissions|CO2|Energy|Demand|Residential and Commercial",
    "Demand - Industry"           = "Gross Emissions|CO2|Energy|Demand|Industry",
    "Industrial Processes"        = "Gross Emissions|CO2|Industrial Processes",
    "Demand - Transport"          = "Emissions|CO2|Energy|Demand|Transportation",
    #"Demand - Other Sector"      = "Emissions|CO2|Energy|Demand|Other",
    "Energy Supply - Other"       = "Gross Emissions|CO2|Energy|Supply|Other",
    "Energy supply - Hydrogen"    = "Gross Emissions|CO2|Energy|Supply|H2",
    "Energy Supply - Solids"      = "Gross Emissions|CO2|Energy|Supply|Solids",
    "Energy Supply - Liquids"     = "Gross Emissions|CO2|Energy|Supply|Liquids",
    "Energy Supply - Gases"       = "Gross Emissions|CO2|Energy|Supply|Gases",
    "Energy Supply - Heat"        = "Gross Emissions|CO2|Energy|Supply|Heat",
    "Energy Supply - Electricity" = "Gross Emissions|CO2|Energy|Supply|Electricity",
    "LULUCF"                      = "Emissions|CO2|AFOLU synthetic",
    "Geological Carbon Removal"   = "Carbon Removal|Geological Storage|ForPlotting"#,
    # "Other Removal"             = "Emissions|CO2|Other Removal"
    )

  totalVars <- c(
    "Total net Kyoto gases"                             = "Emi|Kyoto Gases|synthetic"
    )
  
  varToSector <- setNames(c(names(vars),rev(names(totalVars))), c(vars,rev(totalVars)))

  histVars <- c(
    "Demand - Bunkers"            = c("Emi|CO2|Energy|Demand|Transport|International Bunkers"),
    "non-CO2"                     = c("Emi|Kyoto Gases|non-CO2 calc"),
    "Demand - Buildings"          = c("Emi|CO2|Energy|Demand|Buildings"),
    "Demand - Industry"           = c("Emi|CO2|Energy|Demand|Industry"),
    "Industrial Processes"        = c("Emi|CO2|Industrial Processes"),
    "Demand - Transport"          = c("Emi|CO2|w/o Bunkers|Energy|Demand|Transport"),
    "Energy Supply - Other"       = c("Emi|CO2|Calc Other"),
    "Energy Supply - Solids"      = c("Emi|CO2|Energy|Supply|Solids w/ couple prod"),
    "Energy Supply - Liquids"     = c("Emi|CO2|Energy|Supply|Liquids w/ couple prod"),
    "Energy Supply - Electricity" = c("Emi|CO2|Energy|Supply|Electricity and Heat"),
    "LULUCF"                      = c("Emi|CO2|Land-Use Change|LULUCF national accounting")
    )
  
  totalHistVars <- c(
    "Total net Kyoto gases"                             = "Emi|GHG|w/ Bunkers|LULUCF national accounting"
    )
  
  histVarToSector <- setNames(c(names(histVars),names(totalHistVars)),c(histVars,totalHistVars))

  sector_color <- c(  
    "Demand - Bunkers"            = "#283280",
    "non-CO2"                     = "purple",
    "other CO2"                   = "black",
    "Demand - AFOFI"              = "#6df25e",
    "Demand - Buildings"          = "#eba357",
    "Demand - Industry"           = "#ba6759",
    "Industrial Processes"        = "#735550",
    "Demand - Transport"          = "#a7c8db",
    #"Demand - Other Sector"      = "#5ed5f0",
    "Energy Supply - Other"       = "#ff33ff",
    "Energy supply - Hydrogen"    = "#35effc",
    "Energy Supply - Solids"      = "#0c0c0c",
    "Energy Supply - Liquids"     = "#813499",
    "Energy Supply - Gases"       = "#337fff",
    "Energy Supply - Heat"        = "#b30000",
    "Energy Supply - Electricity" = "#fcec35",
    "LULUCF"                      = "darkgreen",
    "Geological Carbon Removal"   = "red"
    )

  totalShape <- c(
    "Total net Kyoto gases"                             = 17
    )
  
  plotData <- df %>%
    filter(variable %in% c(vars, totalVars),
           period %in% c(2020,2030,2040,2050),
           model %in% modelsFullSystem,
           scenario == "WP1 NetZero",
           region == "EU27 & UK (*)") %>%
    mutate(sector = varToSector[as.character(variable)]) %>%
    rbind(
      histData  %>%
        filter(variable %in% c(histVars, totalHistVars),
               region == "EUR",
               period == 2020,
               model == "UNFCCC") %>%
          mutate(sector = histVarToSector[as.character(variable)],
                 model = "Historical")
      ) %>%
    rbind(
      histData  %>%
        filter(variable %in% c(histVars, totalHistVars),
               region == "EUR",
               period %in% c(2018:2022),
               model == "UNFCCC") %>%
        group_by(model, scenario, region, variable, unit) %>%
        summarize(value = mean(value, na.rm = TRUE)) %>%
        ungroup() %>%
        mutate(sector = histVarToSector[as.character(variable)],
               model = "Historical (5yr average)",
               period = "2020") %>%
        select(all_of(names(df)), sector)
      ) %>%
    mutate(model = factor(model, levels=rev(c(modelsOrder[modelsOrder != "Historical"], "Historical (5yr average)", "Historical"))),
           sector = factor(sector, levels = c(names(vars), names(totalVars))))
    

	g[["Emissions per sector"]] <- ggplot(plotData %>% filter(variable %in% c(vars, histVars))) +
    geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
	  geom_bar(stat='identity', alpha = 0.8,aes(x=model,y=value,fill=sector)) +
    geom_point(data = plotData %>% filter(variable %in% c(totalVars, totalHistVars)), aes(x=model, y = value, shape = sector), colour = "black", size = 3) + 
    facet_wrap(~period, nrow=1, scales = "free_x") +
    theme_paper() +
    scale_fill_manual(values = sector_color, breaks = rev(c("LULUCF", "Geological Carbon Removal", rev(names(sector_color)[!(names(sector_color) %in% c("LULUCF","Geological Carbon Removal"))])))) +
    scale_shape_manual(values = totalShape) +
    theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5)) +
    labs(title = "(a) Emissions per sector", subtitle = expression(paste("Mt ", CO[2],"e /yr"))) +
    theme(legend.key.width = unit(0.7, 'cm'),
          legend.position = "right",
          legend.box = "vertical") + 
    guides(fill = guide_legend(order = 2), 
           shape = guide_legend(order = 1)) + 
    force_panelsizes(cols = plotData %>% select(model, period) %>% group_by(period) %>% summarize(unique_count = n_distinct(model)) %>% pull(unique_count))

```


```{r Carbon capture scale-up, echo=FALSE, include=FALSE}


plotDataRaw <- df %>%
  filter(variable %in% c("Carbon Capture"),
         period >= 2010, period <= 2050,
         scenario %in% c("WP1 NetZero"),
         model %in% modelsFullSystem,
         region == "EU27 & UK (*)")

plotData <- plotDataRaw %>%
  rbind(
    plotDataRaw %>%
      filter(model != "Historical") %>%
      group_by(scenario, region, variable, unit, period) %>%
      summarise(median_value = median(value, na.rm = TRUE)) %>%
      ungroup() %>%
      rename(value = median_value) %>%
      mutate(model = "median") %>%
      select(all_of(names(plotDataRaw)))
  ) %>%
  mutate(model = factor(model, levels=rev(modelsOrder)))

decade_change <- plotData %>% 
  filter(model == "median", period %% 10 == 0, period > 2020) %>%
  select(period, variable, value) %>%
  arrange(variable,period) %>%
  group_by(variable) %>%
  mutate(
    change_in_10_years = value - lag(value),
    change_percentage = ((change_in_10_years) / lag(value))/10*100,
    prev_period = lag(period),
    prev_value = lag(value)
  ) %>%
  mutate(
    x = (period + prev_period) / 2,
    y = (value + prev_value) / 2,
    vAlign = case_when(
      #period == 2050 ~ 1.4,
      TRUE ~ -1.5
    )
  )

g[["Carbon Capture"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_segment(data = plotData %>% filter(period %% 10 == 0), aes(x = period, xend = period, y = 0, yend = value), color = def$grid$color, size = def$grid$lineSize, linetype = def$grid$linetype) + # vertical grid line
  annotate("rect", xmin = -Inf, xmax = histEnd(plotData), ymin = -Inf, ymax = Inf, fill = def$hist$areaColor, alpha = def$hist$alpha) + # historical area
  annotate("text", x = 2017.5, y = 500, label = "Historical", hjust = 0.5, vjust = 0.5, size = 15/.pt, color = "#888888") + # historical text
  geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
  geom_line(data=plotData %>% filter((!(model %in% c("Historical", "median")))), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(!(model %in% c("Historical", "median")), period > 2023), shape = def$model$shape, aes(colour = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) +
  geom_line(data=plotData %>% filter(model == "median"), size=def$median$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(model == "median"), aes(colour = model), shape = def$median$shape, fill = "white", size = def$median$pointSize, stroke = def$median$pointStroke) +
  theme_paper() +
  scale_color_manual(name = "legend", values = setNames(color[levels(plotData$model)],levels(plotData$model)), guide = guide_legend(ncol = 1)) +
  scale_alpha_manual(values = modelAlpha) +
  scale_linetype_manual(values = modelLinetype) +
  labs(title = "(b) Carbon Capture", subtitle = expression(paste("Mt ", CO[2]," /yr"))) + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1000, by = 200), limits = c(0,1050), expand = c(0, 0)) +
  theme(strip.text = element_blank()) +
  coord_cartesian(clip = 'off') +
  theme(legend.position = "right",
        legend.box = "vertical") +
  guides(alpha = "none", linetype = "none")


```


```{r echo=FALSE, include=FALSE}

g[["Figure 5"]] <- gridExtra::arrangeGrob(
  g$`Emissions per sector`,
  g$`Carbon Capture` +
    theme(plot.margin = margin(l = 50, r = 50),  
          legend.margin = margin(l = 100, r = 50))
    ,
  layout_matrix = rbind(c(1),
                        c(1),
                        c(1),
                        c(2),
                        c(2))
   )

ggsave(paste0(outFolder, "/png/Figure 5. Emissions per sector and carbon capture scale-up.png"), g[["Figure 5"]], device="png", width = 12, height = 10, dpi=300, units = "in")
ggsave(paste0(outFolder, "/svg/Figure 5. Emissions per sector and carbon capture scale-up.svg"), g[["Figure 5"]], device="svg", width = 12, height = 10, dpi=100, units = "in")

```


```{r, echo=FALSE, fig.width=12, fig.height=10}

grid.newpage()
grid.draw(g[["Figure 5"]])

```

Figure 5. Emissions per sector and carbon capture scale-up.

---

## Figure 6

<!-- Carbon Price -->

```{r Carbon Price, echo=FALSE, include=FALSE}

plotDataRaw <- df %>%
    filter(variable == "Price|Carbon (weighted by Electricity Demand)",
           period >=2015, period <= 2050,
           model %in% modelsElecall,
           model != "Euro-Calliope",
           scenario %in% c("WP1 NetZero"),
           region == "EU27 & UK (*)")

quartiles_CarbonPrices <- plotDataRaw %>%
  group_by(period) %>%
  summarise_at(vars(value), list( Min = min, Mean = mean, Median = median, Max = max, Sd = sd, Q1=~quantile(., probs = 0.25), Q3=~quantile(., probs = 0.75)))

    quantile(filter(plotData, period == 2030)[["value"]], c(0, 0.25, 0.5, 0.75, 1)) # give out interquartiles    
    quantile(filter(plotData, period == 2040)[["value"]], c(0, 0.25, 0.5, 0.75, 1)) # give out interquartiles     
    quantile(filter(plotData, period == 2050)[["value"]], c(0, 0.25, 0.5, 0.75, 1)) # give out interquartiles 

plotData <- plotDataRaw %>%
  rbind(
    plotDataRaw %>%
      filter(model != "Historical") %>%
      group_by(scenario, region, variable, unit, period) %>%
      summarise(median_value = median(value, na.rm = TRUE)) %>%
      ungroup() %>%
      rename(value = median_value) %>%
      mutate(model = "median") %>%
      select(all_of(names(plotDataRaw)))
  ) %>%
  mutate(model = factor(model, levels=rev(modelsOrder)))

decade_change <- plotData %>% 
  filter(model == "median", period %% 10 == 0, period > 2020) %>%
  select(period, variable, value) %>%
  arrange(variable,period) %>%
  group_by(variable) %>%
  mutate(
    change_in_10_years = value - lag(value),
    change_percentage = ((change_in_10_years) / lag(value))/10 * 100,
    prev_period = lag(period),
    prev_value = lag(value)
  ) %>%
  mutate(
    x = (period + prev_period) / 2,
    y = (value + prev_value) / 2,
    vAlign = case_when(
      TRUE ~ -1.5
    )
  )

g[["Carbon Price"]] <- ggplot(plotData,aes(x=period,y=value)) +
  geom_segment(data = plotData %>% filter(period %% 10 == 0), aes(x = period, xend = period, y = 0, yend = value), color = def$grid$color, size = def$grid$lineSize, linetype = def$grid$linetype) + # vertical grid line
  annotate("rect", xmin = -Inf, xmax = histEnd(plotData), ymin = -Inf, ymax = Inf, fill = def$hist$areaColor, alpha = def$hist$alpha) + # historical area
  annotate("text", x = 2019.5, y = 600, label = "Historical", hjust = 0.5, vjust = 0.5, size = 15/.pt, color = "#888888") + # historical text
  geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
  geom_line(data=plotData %>% filter((!(model %in% c("Historical", "median")))), size=def$model$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(model != "median", period > 2023), shape = def$model$shape, aes(colour = model, shape = model), fill = "white", size = def$model$pointSize, stroke = def$model$pointStroke) +
  geom_line(data=plotData %>% filter(model == "median"), size=def$median$lineSize,aes(color=model,linetype=model, alpha = model)) + # models data
  geom_point(data = plotData %>% filter(model == "median"), shape = def$median$shape, aes(colour = model, shape = model), fill = "white", size = def$median$pointSize, stroke = def$median$pointStroke) +
  theme_paper() +
  scale_color_manual(name = "legend", values = setNames(color[levels(plotData$model)],levels(plotData$model)), guide = guide_legend(nrow = 2)) +
  scale_alpha_manual(values = modelAlpha) +
  scale_linetype_manual(values = modelLinetype) +
  scale_shape_manual(name = "legend", values = modelShape, guide = guide_legend(ncol = 1)) +
  labs(title = expression(paste("(a) Carbon Price")), subtitle = expression(paste("€/t", CO[2]))) + 
  scale_x_continuous(expand = c(0, 0)) +
  coord_cartesian(clip = 'off') + 
  theme(legend.position = "bottom",
        legend.key.height = unit(1, 'cm'),
        plot.margin = margin(r = 50, l = 20),
        #legend.margin = margin(l = 50),
        strip.text = element_blank()
  ) +
  guides(alpha = "none", linetype = "none")

```


```{r Mitigation costs, echo=FALSE, include=FALSE}

  plotDataRaw <- df %>%
    filter(variable == "Mitigation costs",
           period == 2050,
           scenario %in% c("WP1 NetZero"),
           region == "EU27 & UK (*)")

  plotData <- plotDataRaw %>%
    mutate(
      min = min(value),
      max = max(value))

  g[["Mitigation costs"]] <- ggplot(plotData, aes(x = 0, y = value)) +
    geom_hline(yintercept=0, color = "black", linewidth= def$yAxis$linewidth, alpha = def$yAxis$alpha) + # y-axis zero
    geom_errorbar(aes(ymin=min, ymax=max), width=0.3, colour="black", alpha=0.9, linewidth=1) +
    geom_boxplot(outlier.shape = NA, width=0.3) +
    geom_point(data = plotData, shape = def$model$shape, aes(colour = model), fill = "white", size = 2, stroke = 2) +
    geom_text_repel(aes(label = model, colour = model),
      nudge_x = c(0.3,0.3,0.3,0.3,0.3),
      nudge_y = c(-0.001,-0.001,0.001,0.002,0.001),
      direction = "y",
      hjust = 0,
      segment.size = 0.6,
      segment.color = "grey30",
      size = 3
    ) +
    theme_paper() +
    scale_y_continuous(labels = scales::percent_format(accuracy = 0.1), breaks = seq(0, 0.025, by = 0.005), limits = c(0,0.025)) +
    scale_x_continuous(limits = c(-0.6, 0.6)) + #, expand = expansion(mult = 0.2)) +
    theme(
      legend.position = "none",
      axis.text.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      axis.ticks.x = element_blank(),
      plot.margin = margin(b = 100, r = 20),
    ) +
    labs(title = expression(paste("(b) Mitigation costs")), subtitle = "as share of consumption")

  
```


```{r figure 6, echo=FALSE, include=FALSE}

g[["Figure 6"]] <- gridExtra::arrangeGrob(
  g$`Carbon Price`,
  g$`Mitigation costs`,
  layout_matrix = rbind(c(1,1,1,2))
   )

ggsave(paste0(outFolder, "/png/Figure 6 Carbon price and Mitigation costs.png"), g[["Figure 6"]], device="png", width = 12, height = 6, dpi=300, units = "in")
ggsave(paste0(outFolder, "/svg/Figure 6 Carbon price and Mitigation costs.svg"), g[["Figure 6"]], device="svg", width = 12, height = 6, dpi=100, units = "in")

```


```{r, echo=FALSE, fig.width=12, fig.height=6}

grid.newpage()
grid.draw(g[["Figure 6"]])

```

Figure 6. Carbon price and Share of consumption. 


---

## Additional Stacked Bar Charts

```{r stacked_helpers, echo=FALSE, include=FALSE}

stackedModels     <- c("Historical", "Historical (5yr average)", modelsFullSystem)
stackedElecModels <- c(modelsFullSystem, "Euro-Calliope", "LIMES", "MEESA", "OSeMBE")

# Two historical bars for the stacked charts - the 2020 value and the 2018-2022 average
histStackedBars <- function(map, histModel = "IEA") {
  if (!("weight" %in% names(map))) map$weight <- 1
  if (!("source" %in% names(map))) map$source <- histModel
  map$source[is.na(map$source)] <- histModel
  groupCols <- setdiff(names(map), c("variable", "weight", "source"))

  perYear <- histData %>%
    filter(region == "EUR", period %in% 2018:2022) %>%
    mutate(source = as.character(model), variable = as.character(variable)) %>%
    select(source, variable, period, value) %>%
    inner_join(map, by = c("source", "variable")) %>%
    mutate(value = value * weight) %>%
    group_by(across(all_of(c(groupCols, "period")))) %>%
    summarise(value = sum(value, na.rm = TRUE), .groups = "drop")

  bind_rows(
    perYear %>%
      filter(period == 2020) %>%
      mutate(model = "Historical"),
    perYear %>%
      group_by(across(all_of(groupCols))) %>%
      summarise(value = mean(value, na.rm = TRUE), .groups = "drop") %>%
      mutate(model = "Historical (5yr average)", period = 2020)
  )
}

# Number of bars per facet, so the 2020 panel can be widened to fit the historical bars.
panelBarCounts <- function(...) {
  bind_rows(lapply(list(...), function(x) x %>% select(model, period))) %>%
    group_by(period) %>%
    summarise(n = n_distinct(model), .groups = "drop") %>%
    arrange(period) %>%
    pull(n)
}

```

### Primary Energy

Historical source: IEA data
- IEA covers every category over 1990-2023. Nuclear, hydro, wind and solar are only reported at the carrier level, hence the |Electricity variables; geothermal is split into an electricity and a heat variable which stack into a single segment.
- The IEA reports nuclear as primary heat, the models as direct equivalent, so nuclear is scaled by 0.334 in both the bars and the total.
- The IEA primary energy total also covers waste and heat, to match the total to the bars, the difference is added to "Other".

```{r primary_energy_stacked, echo=FALSE, include=FALSE}

pe_vars <- c(
  "Coal"       = "Primary Energy|Coal",
  "Oil"        = "Primary Energy|Oil",
  "Gas"        = "Primary Energy|Gas",
  "Biomass"    = "Primary Energy|Biomass",
  "Hydro"      = "Primary Energy|Hydro",
  "Wind"       = "Primary Energy|Wind",
  "Solar"      = "Primary Energy|Solar",
  "Nuclear"    = "Primary Energy|Nuclear",
  "Geothermal" = "Primary Energy|Geothermal",
  "Other"      = "Primary Energy|Other"
)

pe_colors <- c(
  "Coal"       = "#0c0c0c",
  "Oil"        = "#813499",
  "Gas"        = "#337fff",
  "Biomass"    = "#005900",
  "Hydro"      = "#0000cc",
  "Wind"       = "#35effc",
  "Solar"      = "#fcec35",
  "Nuclear"    = "#ff33ff",
  "Geothermal" = "#b30000",
  "Other"      = "#79a6d2"
)

# IEA covers every category over 1990-2023. Nuclear, hydro, wind and solar are only
# reported at the carrier level, hence the |Electricity variables; geothermal is split
# into an electricity and a heat variable which stack into a single segment.
pe_hist_map <- data.frame(
  variable = c("PE|Coal", "PE|Oil", "PE|Gas", "PE|Biomass", "PE|Hydro|Electricity",
               "PE|Wind|Electricity", "PE|Solar|Electricity", "PE|Uranium|Electricity",
               "PE|Geothermal|Electricity", "PE|Geothermal|Heat"),
  category = c("Coal", "Oil", "Gas", "Biomass", "Hydro",
               "Wind", "Solar", "Nuclear",
               "Geothermal", "Geothermal"),
  weight   = c(1, 1, 1, 1, 1,
               1, 1, 0.334,
               1, 1),
  stringsAsFactors = FALSE
)

# The IEA primary energy total also covers waste and heat, to match the total to the bars, the difference is added to "Other".
pe_hist_map <- bind_rows(
  pe_hist_map,
  data.frame(variable = "PE", category = "Other", weight = 1, stringsAsFactors = FALSE),
  pe_hist_map %>% mutate(category = "Other", weight = -1)
)

plotData_PE <- df %>%
  filter(
    variable %in% pe_vars,
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsFullSystem
  )

if ("EU27 & UK (*)" %in% unique(plotData_PE$region)) {
  plotData_PE <- plotData_PE %>% filter(region == "EU27 & UK (*)")
}

peLabelMap <- setNames(names(pe_vars), pe_vars)
plotData_PE <- plotData_PE %>%
  mutate(model = as.character(model), category = peLabelMap[as.character(variable)]) %>%
  select(model, period, category, value) %>%
  bind_rows(histStackedBars(pe_hist_map)) %>%
  mutate(category = factor(category, levels = names(pe_vars)),
         model = factor(model, levels = stackedModels))

plotData_PE_Total <- df %>%
  filter(
    variable %in% c("Primary Energy", "Primary Energy|w/o Bunkers"),
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsFullSystem
  )
if ("EU27 & UK (*)" %in% unique(plotData_PE_Total$region)) {
  plotData_PE_Total <- plotData_PE_Total %>% filter(region == "EU27 & UK (*)")
}

pe_hist_total_map <- data.frame(
  variable = c("PE", "PE|Uranium|Electricity"),
  weight   = c(1, 0.334 - 1),
  stringsAsFactors = FALSE
)

plotData_PE_Total <- plotData_PE_Total %>%
  mutate(model = as.character(model)) %>%
  select(model, period, value) %>%
  bind_rows(histStackedBars(pe_hist_total_map)) %>%
  mutate(model = factor(model, levels = stackedModels))

g[["primary_energy_stacked"]] <- ggplot(plotData_PE) +
  geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
  geom_bar(aes(x = model, y = value, fill = category), stat = 'identity', alpha = 0.8) +
  geom_point(data = plotData_PE_Total, aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
  theme_paper() +
  facet_wrap(~period, nrow = 1, scales = "free_x") +
  scale_fill_manual(values = pe_colors) +
  coord_cartesian(clip = 'off') +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
    strip.text = element_text(size = 12, colour = "black"),
    plot.margin = margin(t = 15, r = 25, b = 15, l = 20)
  ) +
  ylab("Primary Energy (EJ/yr)") +
  labs(title = "Primary Energy by Carrier") +
  force_panelsizes(cols = panelBarCounts(plotData_PE, plotData_PE_Total))

ggsave(paste0(outFolder, "/svg/primary_energy_stacked.svg"), g[["primary_energy_stacked"]], device = "svg", width = 12, height = 8, dpi = 100, units = "in")
ggsave(paste0(outFolder, "/png/primary_energy_stacked.png"), g[["primary_energy_stacked"]], device = "png", width = 12, height = 8, dpi = 100, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["primary_energy_stacked"]]

```

---

### Carbon Capture by Origin

```{r carbon_capture_origin_stacked, echo=FALSE, include=FALSE}

cc_vars <- c(
  "Biomass (BECCS)"      = "Carbon Capture|Biomass",
  "DACCS"                = "Carbon Capture|Direct Air Capture",
  "Fossil"               = "Carbon Capture|Fossil",
  "Industrial Processes" = "Carbon Capture|Industrial Processes",
  "Other"                = "Carbon Capture|Other"
)

cc_colors <- c(
  "Biomass (BECCS)"      = "#005900",
  "DACCS"                = "#CC6666",
  "Fossil"               = "#813499",
  "Industrial Processes" = "#5c5c5c",
  "Other"                = "#79a6d2"
)

plotData_CC <- df %>%
  filter(
    variable %in% cc_vars,
    period %in% c(2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% c(modelsFullSystem, "Euro-Calliope")
  )

if ("EU27 & UK (*)" %in% unique(plotData_CC$region)) {
  plotData_CC <- plotData_CC %>% filter(region == "EU27 & UK (*)")
}

if (nrow(plotData_CC) > 0) {
  ccLabelMap <- setNames(names(cc_vars), cc_vars)
  plotData_CC$variable <- factor(ccLabelMap[as.character(plotData_CC$variable)], levels = names(cc_vars))
  plotData_CC$model <- factor(as.character(plotData_CC$model), levels = c(modelsFullSystem, "Euro-Calliope"))

  plotData_CC_Total <- df %>%
    filter(
      variable %in% c("Carbon Capture"), #, "Carbon Sequestration", "Carbon Capture|Storage"
      period %in% c(2030, 2040, 2050),
      scenario == "WP1 NetZero",
      model %in% c(modelsFullSystem, "Euro-Calliope")
    )
  if ("EU27 & UK (*)" %in% unique(plotData_CC_Total$region)) {
    plotData_CC_Total <- plotData_CC_Total %>% filter(region == "EU27 & UK (*)")
  }
  plotData_CC_Total$model <- factor(as.character(plotData_CC_Total$model), levels = levels(plotData_CC$model))

  g[["carbon_capture_origin_stacked"]] <- ggplot(plotData_CC %>% filter(model != "Euro-Calliope")) +
    geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
    geom_bar(aes(x = model, y = value, fill = variable), stat = 'identity', alpha = 0.8) +
    geom_point(data = plotData_CC_Total %>% filter(model != "Euro-Calliope"), aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
    theme_paper() +
    facet_wrap(~period, nrow = 1) +
    scale_fill_manual(values = cc_colors) +
    coord_cartesian(clip = 'off') +
    theme(
      axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
      strip.text = element_text(size = 12, colour = "black"),
      plot.margin = margin(t = 15, r = 25, b = 15, l = 20)
    ) +
    ylab("Carbon Capture (Mt CO2/yr)") +
    labs(title = "Carbon Capture by Origin")
  
  g[["carbon_capture_origin_stacked_with_EuroCalliope"]] <- ggplot(plotData_CC) +
    geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
    geom_bar(aes(x = model, y = value, fill = variable), stat = 'identity', alpha = 0.8) +
    geom_point(data = plotData_CC_Total, aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
    theme_paper() +
    facet_wrap(~period, nrow = 1) +
    scale_fill_manual(values = cc_colors) +
    coord_cartesian(clip = 'off') +
    theme(
      axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
      strip.text = element_text(size = 12, colour = "black"),
      plot.margin = margin(t = 15, r = 25, b = 15, l = 20)
    ) +
    ylab("Carbon Capture (Mt CO2/yr)") +
    labs(title = "Carbon Capture by Origin")

  ggsave(paste0(outFolder, "/svg/carbon_capture_origin_stacked.svg"), g[["carbon_capture_origin_stacked"]], device = "svg", width = 12, height = 8, dpi = 100, units = "in")
  ggsave(paste0(outFolder, "/png/carbon_capture_origin_stacked.png"), g[["carbon_capture_origin_stacked"]], device = "png", width = 12, height = 8, dpi = 100, units = "in")
  ggsave(paste0(outFolder, "/svg/carbon_capture_origin_stacked_with_EuroCalliope.svg"), g[["carbon_capture_origin_stacked_with_EuroCalliope"]], device = "svg", width = 12, height = 8, dpi = 100, units = "in")
  ggsave(paste0(outFolder, "/png/carbon_capture_origin_stacked_with_EuroCalliope.png"), g[["carbon_capture_origin_stacked_with_EuroCalliope"]], device = "png", width = 12, height = 8, dpi = 100, units = "in")
}

```

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["carbon_capture_origin_stacked"]]

```

---

### Final Energy by Carrier

Historical source: IEA data 
- The IEA carriers add up exactly to the IEA final energy total, bunkers included, which matches the models "Final Energy" used here (checked again on the current historical.mif: agrees to 1e-4 for 2018-2023).

```{r final_energy_carrier_stacked, echo=FALSE, include=FALSE}

fe_carrier_vars <- c(
  "Hydrogen"    = "Final Energy|Hydrogen",
  "Gases"       = "Final Energy|Gases",
  "Liquids"     = "Final Energy|Liquids",
  "Solids"      = "Final Energy|Solids",
  "Other"       = "Final Energy|Other",
  "Heat"        = "Final Energy|Heat",
  "Electricity" = "Final Energy|Electricity"
)

fe_carrier_colors <- c(
  "Hydrogen"    = "#5ed5b0",
  "Gases"       = "#999959",
  "Liquids"     = "#0000cc",
  "Solids"      = "#0c0c0c",
  "Other"       = "#79a6d2",
  "Heat"        = "#cc0000",
  "Electricity" = "#ffb200"
)

# The IEA carriers add up exactly to the IEA final energy total including bunkers in the historical.mif.
fe_carrier_hist_map <- data.frame(
  variable = c("FE|Hydrogen", "FE|Gases", "FE|Liquids", "FE|Solids", "FE|Heat", "FE|Electricity"),
  category = c("Hydrogen", "Gases", "Liquids", "Solids", "Heat", "Electricity"),
  stringsAsFactors = FALSE
)

plotData_FEC <- df %>%
  filter(
    variable %in% fe_carrier_vars,
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsFullSystem
  )

if ("EU27 & UK (*)" %in% unique(plotData_FEC$region)) {
  plotData_FEC <- plotData_FEC %>% filter(region == "EU27 & UK (*)")
}

fecLabelMap <- setNames(names(fe_carrier_vars), fe_carrier_vars)
plotData_FEC <- plotData_FEC %>%
  mutate(model = as.character(model), category = fecLabelMap[as.character(variable)]) %>%
  select(model, period, category, value) %>%
  bind_rows(histStackedBars(fe_carrier_hist_map)) %>%
  mutate(category = factor(category, levels = names(fe_carrier_vars)),
         model = factor(model, levels = stackedModels))

plotData_FEC_Total <- df %>%
  filter(
    variable == "Final Energy",
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsFullSystem
  )
if ("EU27 & UK (*)" %in% unique(plotData_FEC_Total$region)) {
  plotData_FEC_Total <- plotData_FEC_Total %>% filter(region == "EU27 & UK (*)")
}
plotData_FEC_Total <- plotData_FEC_Total %>%
  mutate(model = as.character(model)) %>%
  select(model, period, value) %>%
  bind_rows(histStackedBars(data.frame(variable = "FE", stringsAsFactors = FALSE))) %>%
  mutate(model = factor(model, levels = stackedModels))

g[["final_energy_carrier_stacked"]] <- ggplot(plotData_FEC) +
  geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
  geom_bar(aes(x = model, y = value, fill = category), stat = 'identity', alpha = 0.8) +
  geom_point(data = plotData_FEC_Total, aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
  theme_paper() +
  facet_wrap(~period, nrow = 1, scales = "free_x") +
  scale_fill_manual(values = fe_carrier_colors) +
  coord_cartesian(clip = 'off') +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
    strip.text = element_text(size = 12, colour = "black"),
    plot.margin = margin(t = 15, r = 25, b = 15, l = 20)
  ) +
  ylab("Final Energy (EJ/yr)") +
  labs(title = "Final Energy by Carrier") +
  force_panelsizes(cols = panelBarCounts(plotData_FEC, plotData_FEC_Total))

ggsave(paste0(outFolder, "/svg/final_energy_carrier_stacked.svg"), g[["final_energy_carrier_stacked"]], device = "svg", width = 12, height = 8, dpi = 100, units = "in")
ggsave(paste0(outFolder, "/png/final_energy_carrier_stacked.png"), g[["final_energy_carrier_stacked"]], device = "png", width = 12, height = 8, dpi = 100, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["final_energy_carrier_stacked"]]

```

---

### Final Energy per Sector

Historical source: IEA data 
- The IEA has `FE = FE|Industry + FE|Transport + FE|Buildings`, so non-energy is inside industry and bunkers inside transport. Industry comes from the IEA's own `FE|w/o Non-energy Use|Industry`; bunkers I still subtract, to match the models "Final Energy (w/o bunkers)".
- The IEA has no agriculture or other-sector split.

```{r final_energy_sector_stacked, echo=FALSE, include=FALSE}

fe_sector_vars <- c(
  "Industry"                 = "Final Energy|Industry",
  "Non-Energy Use"           = "Final Energy|Non-Energy Use",
  "Transportation"           = "Final Energy|Transportation",
  "Residential & Commercial" = "Final Energy|Residential and Commercial",
  "Agriculture"              = "Final Energy|Agriculture",
  "Other Sector"             = "Final Energy|Other Sector"
)

if (!("Final Energy|Transportation" %in% unique(df$variable)) && ("Final Energy|Transport" %in% unique(df$variable))) {
  fe_sector_vars["Transportation"] <- "Final Energy|Transport"
}
if (!("Final Energy|Residential and Commercial" %in% unique(df$variable)) && ("Final Energy|Buildings" %in% unique(df$variable))) {
  fe_sector_vars["Residential & Commercial"] <- "Final Energy|Buildings"
}

fe_sector_colors <- c(
  "Industry"                 = "#5f6769",
  "Non-Energy Use"           = "#cccccc",
  "Transportation"           = "#a1dd70",
  "Residential & Commercial" = "#5edfff",
  "Agriculture"              = "#005900",
  "Other Sector"             = "#999959"
)

# The IEA has `FE = FE|Industry + FE|Transport + FE|Buildings`, so non-energy is inside industry and bunkers inside transport.
# Industry comes from the IEA's own w/o-non-energy variable; bunkers I still subtract, to match the models "Final Energy (w/o bunkers)".
# The IEA has no agriculture or other-sector split.
fe_sector_hist_map <- data.frame(
  variable = c("FE|w/o Non-energy Use|Industry", "FE|Non-energy Use", "FE|Transport",
               "FE|Transport|Bunkers", "FE|Buildings", "FE|AFOFI", "FE|Other Sector"),
  category = c("Industry", "Non-Energy Use", "Transportation",
               "Transportation", "Residential & Commercial", "Agriculture", "Other Sector"),
  weight   = c(1, 1, 1, -1, 1, 1, 1),
  stringsAsFactors = FALSE
)

# Matching total: the IEA total includes bunkers, the models' reference does not.
fe_sector_hist_total_map <- data.frame(
  variable = c("FE", "FE|Transport|Bunkers"),
  weight   = c(1, -1),
  stringsAsFactors = FALSE
)

plotData_FES <- df %>%
  filter(
    variable %in% fe_sector_vars,
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsFullSystem
  )

if ("EU27 & UK (*)" %in% unique(plotData_FES$region)) {
  plotData_FES <- plotData_FES %>% filter(region == "EU27 & UK (*)")
}

fesLabelMap <- setNames(names(fe_sector_vars), fe_sector_vars)
plotData_FES <- plotData_FES %>%
  mutate(model = as.character(model), category = fesLabelMap[as.character(variable)]) %>%
  select(model, period, category, value) %>%
  bind_rows(histStackedBars(fe_sector_hist_map)) %>%
  mutate(category = factor(category, levels = names(fe_sector_vars)),
         model = factor(model, levels = stackedModels))

plotData_FES_Total <- df %>%
  filter(
    variable == "Final Energy (w/o bunkers)",
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsFullSystem
  )
if ("EU27 & UK (*)" %in% unique(plotData_FES_Total$region)) {
  plotData_FES_Total <- plotData_FES_Total %>% filter(region == "EU27 & UK (*)")
}
plotData_FES_Total <- plotData_FES_Total %>%
  mutate(model = as.character(model)) %>%
  select(model, period, value) %>%
  bind_rows(histStackedBars(fe_sector_hist_total_map)) %>%
  mutate(model = factor(model, levels = stackedModels))

g[["final_energy_sector_stacked"]] <- ggplot(plotData_FES) +
  geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
  geom_bar(aes(x = model, y = value, fill = category), stat = 'identity', alpha = 0.8) +
  geom_point(data = plotData_FES_Total, aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
  theme_paper() +
  facet_wrap(~period, nrow = 1, scales = "free_x") +
  scale_fill_manual(values = fe_sector_colors) +
  coord_cartesian(clip = 'off') +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
    strip.text = element_text(size = 12, colour = "black"),
    plot.margin = margin(t = 15, r = 25, b = 15, l = 20)
  ) +
  ylab("Final Energy (EJ/yr)") +
  labs(title = "Final Energy per Sector") +
  force_panelsizes(cols = panelBarCounts(plotData_FES, plotData_FES_Total))

ggsave(paste0(outFolder, "/svg/final_energy_sector_stacked.svg"), g[["final_energy_sector_stacked"]], device = "svg", width = 12, height = 8, dpi = 100, units = "in")
ggsave(paste0(outFolder, "/png/final_energy_sector_stacked.png"), g[["final_energy_sector_stacked"]], device = "png", width = 12, height = 8, dpi = 100, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["final_energy_sector_stacked"]]

```

---

### Final Energy per Sector and Carrier

Historical source: IEA data 
- The IEA only splits industry, buildings and transport by carrier, so the historical bars stay empty in the agriculture and other-sector rows.
- The IEA nests non-energy use inside industry, whereas the models report it as its own sector. The sector *totals* need no assumption - the IEA reports `FE|w/o Non-energy Use|Industry` directly - but the *carrier* split still does: the IEA series used here has no non-energy-use breakdown by carrier, so for simplicity I assumed it all feedstock liquids and moved it out of industry liquids. (The current mif does have `FE|Non-energy Use|Industry|Liquids|Fossil`, but only on the IEA WEO 2025 rows, not the IEA series used here.) Be aware that this causes only approximated comparisons on gas and liquids carrier levels at industry and non-energy use that I would avoid/fix before showing in any publication.


```{r final_energy_sector_carrier_stacked, echo=FALSE, include=FALSE}

fe_carrier_colors <- c(
  "Electricity" = "#fcec35",
  "Heat"        = "#b30000",
  "Hydrogen"    = "#00ccff",
  "Gases"       = "#337fff",
  "Liquids"     = "#813499",
  "Solids"      = "#0c0c0c",
  "Other"       = "#79a6d2"
)

fe_sc_map <- expand.grid(
  sector_raw = c("Industry", "Non-Energy Use","Transportation", "Transport", "Residential and Commercial", "Buildings", "Agriculture", "Other Sector"),
  carrier_raw = c("Electricity", "Gases", "Heat", "Hydrogen", "Liquids", "Solids", "Other"),
  stringsAsFactors = FALSE
) %>%
  mutate(
    variable = paste0("Final Energy|", sector_raw, "|", carrier_raw),
    sector = case_when(
      sector_raw %in% c("Transportation", "Transport") ~ "Transportation",
      sector_raw %in% c("Residential and Commercial", "Buildings") ~ "Residential & Commercial",
      TRUE ~ sector_raw
    ),
    carrier = carrier_raw
  )

# The IEA only splits industry, buildings and transport by carrier, so the historical
# bars stay empty in the agriculture and other-sector rows.
fe_sc_hist_map <- expand.grid(
  sector_raw  = c("Industry", "Buildings", "Transport"),
  carrier_raw = c("Electricity", "Gases", "Heat", "Hydrogen", "Liquids", "Solids"),
  stringsAsFactors = FALSE
) %>%
  mutate(
    variable = paste0("FE|", sector_raw, "|", carrier_raw),
    sector = case_when(
      sector_raw == "Transport" ~ "Transportation",
      sector_raw == "Buildings" ~ "Residential & Commercial",
      TRUE ~ sector_raw
    ),
    carrier = carrier_raw,
    weight = 1
  ) %>%
  select(variable, sector, carrier, weight) %>%
  # The IEA nests non-energy use inside industry, whereas the models report it as its own sector. 
  # I couldn't find a historical breakdown of non-energy use by carrier, so for simplicit I assumed it as feedstock liquids and subtracted from industry to match the historical bars totals. 
  # Be aware that this causes only approximated comparisons on gas and liquids carrier levels at industry and non-energy use that I would avoid/fix before showing in any publication.
  bind_rows(
    data.frame(variable = "FE|Non-energy Use", sector = "Industry",
               carrier = "Liquids", weight = -1, stringsAsFactors = FALSE),
    data.frame(variable = "FE|Non-energy Use", sector = "Non-Energy Use",
               carrier = "Liquids", weight = 1, stringsAsFactors = FALSE)
  )

# The sector totals need no subtraction: the IEA reports industry excluding non-energy use
# directly. Only the carrier split above still rests on the all-liquids assumption.
fe_sc_hist_total_map <- data.frame(
  variable = c("FE|w/o Non-energy Use|Industry", "FE|Buildings", "FE|Transport", "FE|Non-energy Use"),
  sector   = c("Industry", "Residential & Commercial", "Transportation", "Non-Energy Use"),
  stringsAsFactors = FALSE
)

plotData_FE_SC <- df %>%
  filter(
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsFullSystem
  ) %>%
  inner_join(fe_sc_map, by = "variable")

if ("EU27 & UK (*)" %in% unique(plotData_FE_SC$region)) {
  plotData_FE_SC <- plotData_FE_SC %>% filter(region == "EU27 & UK (*)")
}

if (nrow(plotData_FE_SC) > 0) {
  plotData_FE_SC <- plotData_FE_SC %>%
    mutate(model = as.character(model)) %>%
    select(model, period, sector, carrier, value) %>%
    bind_rows(histStackedBars(fe_sc_hist_map)) %>%
    mutate(
      sector = factor(sector, levels = c("Industry", "Non-Energy Use", "Transportation", "Residential & Commercial", "Agriculture", "Other Sector")),
      carrier = factor(carrier, levels = names(fe_carrier_colors)),
      model = factor(model, levels = stackedModels)
    )

  fes_total_map <- c(
    "Industry"                 = "Final Energy|Industry",
    "Non-Energy Use"           = "Final Energy|Non-Energy Use",
    "Transportation"           = "Final Energy|Transportation",
    "Residential & Commercial" = "Final Energy|Residential and Commercial",
    "Agriculture"              = "Final Energy|Agriculture",
    "Other Sector"             = "Final Energy|Other Sector"
  )
  if (!("Final Energy|Transportation" %in% unique(df$variable)) && ("Final Energy|Transport" %in% unique(df$variable))) {
    fes_total_map["Transportation"] <- "Final Energy|Transport"
  }
  if (!("Final Energy|Residential and Commercial" %in% unique(df$variable)) && ("Final Energy|Buildings" %in% unique(df$variable))) {
    fes_total_map["Residential & Commercial"] <- "Final Energy|Buildings"
  }

  plotData_FE_SC_Total <- df %>%
    filter(
      variable %in% fes_total_map,
      period %in% c(2020, 2030, 2040, 2050),
      scenario == "WP1 NetZero",
      model %in% modelsFullSystem
    )
  if ("EU27 & UK (*)" %in% unique(plotData_FE_SC_Total$region)) {
    plotData_FE_SC_Total <- plotData_FE_SC_Total %>% filter(region == "EU27 & UK (*)")
  }

  fesTotalLabelMap <- setNames(names(fes_total_map), fes_total_map)
  plotData_FE_SC_Total <- plotData_FE_SC_Total %>%
    mutate(model = as.character(model), sector = fesTotalLabelMap[as.character(variable)]) %>%
    select(model, period, sector, value) %>%
    bind_rows(histStackedBars(fe_sc_hist_total_map)) %>%
    mutate(sector = factor(sector, levels = levels(plotData_FE_SC$sector)),
           model = factor(model, levels = stackedModels))

  g[["final_energy_sector_carrier_stacked"]] <- ggplot(plotData_FE_SC) +
    geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
    geom_bar(aes(x = model, y = value, fill = carrier), stat = 'identity', alpha = 0.8) +
    geom_point(data = plotData_FE_SC_Total, aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
    theme_paper() +
    facet_grid(sector ~ period, scales = "free") +
    scale_fill_manual(values = fe_carrier_colors) +
    coord_cartesian(clip = 'off') +
    theme(
      axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
      strip.text = element_text(size = 11, colour = "black", face = "bold"),
      strip.text.y = element_text(angle = 0, hjust = 0, size = 11, colour = "black", face = "bold"),
      plot.margin = margin(t = 20, r = 40, b = 20, l = 20)
    ) +
    ylab("Final Energy (EJ/yr)") +
    labs(title = "Final Energy per Sector and Carrier") +
    force_panelsizes(cols = panelBarCounts(plotData_FE_SC, plotData_FE_SC_Total))

  ggsave(paste0(outFolder, "/svg/final_energy_sector_carrier_stacked.svg"), g[["final_energy_sector_carrier_stacked"]], device = "svg", width = 14, height = 14, dpi = 100, units = "in")
  ggsave(paste0(outFolder, "/png/final_energy_sector_carrier_stacked.png"), g[["final_energy_sector_carrier_stacked"]], device = "png", width = 14, height = 14, dpi = 100, units = "in")
}

```

```{r, echo=FALSE, fig.width=14, fig.height=14}

if (!is.null(g[["final_energy_sector_carrier_stacked"]])) {
  g[["final_energy_sector_carrier_stacked"]]
}

```

---

### Secondary Energy | Electricity


Historical source: Ember, plus BP for oil
- unlike the IEA, Ember reports the generation total.
- Ember has no oil; BP does, on the same basis (the two agree to within 1% on nuclear, wind, solar and hydro), so oil comes from BP. Not from the IEA: its `SE|Electricity` rows in this mif are labelled EJ/yr but sum to ~127 EJ, about twice the EU's entire primary energy, so they are unusable.
- Geothermal and other generation are still missing (no usable source), so the historical stack sits slightly below its own total.


```{r secondary_energy_electricity_stacked, echo=FALSE, include=FALSE}

elec_colors <- c(
  "Coal"       = "#0c0c0c",
  "Oil"        = "#813499",
  "Gas"        = "#337fff",
  "Biomass"    = "#005900",
  "Nuclear"    = "#ff33ff",
  "Hydro"      = "#0000cc",
  "Wind"       = "#35effc",
  "Solar"      = "#fcec35",
  "Geothermal" = "#b30000",
  "Hydrogen"   = "#5ed5b0",
  "Other"      = "#79a6d2"
)

se_elec_vars <- setNames(paste0("Secondary Energy|Electricity|", names(elec_colors)), names(elec_colors))

# Ember is used for the historical electricity bars rather than the IEA of the other
# charts: it is the reference dataset for EU power and, unlike the IEA, it reports the
# generation total. It has no geothermal or other category, so the historical stack still
# sits slightly below its own total.
se_elec_hist_map <- bind_rows(
  data.frame(
    variable = paste0("SE|Electricity|", c("Coal", "Gas", "Biomass", "Nuclear", "Hydro", "Wind", "Solar")),
    category = c("Coal", "Gas", "Biomass", "Nuclear", "Hydro", "Wind", "Solar"),
    source   = "Ember",
    stringsAsFactors = FALSE
  ),
  # Ember reports no oil; BP does, on the same basis (the two agree to within 1% on
  # nuclear, wind, solar and hydro), so the oil segment comes from BP. NOT from the IEA:
  # its SE|Electricity rows in this mif are labelled EJ/yr but sum to ~127 EJ, about
  # twice the EU's entire primary energy, so they are unusable here.
  data.frame(variable = "SE|Electricity|Oil", category = "Oil", source = "BP",
             stringsAsFactors = FALSE)
)

elecModelOrder <- c("Historical", "Historical (5yr average)", stackedElecModels)
orderElecModels <- function(lims) elecModelOrder[elecModelOrder %in% lims]

plotData_SEE <- df %>%
  filter(
    variable %in% se_elec_vars,
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsElecall
  )

if ("EU27 & UK (*)" %in% unique(plotData_SEE$region)) {
  plotData_SEE <- plotData_SEE %>% filter(region == "EU27 & UK (*)")
}

seeLabelMap <- setNames(names(se_elec_vars), se_elec_vars)
plotData_SEE <- plotData_SEE %>%
  mutate(model = as.character(model), category = seeLabelMap[as.character(variable)]) %>%
  select(model, period, category, value) %>%
  bind_rows(histStackedBars(se_elec_hist_map, histModel = "Ember")) %>%
  mutate(category = factor(category, levels = names(elec_colors)),
         model = factor(model, levels = elecModelOrder))

plotData_SEE_Total <- df %>%
  filter(
    variable == "Secondary Energy|Electricity",
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsElecall
  )
if ("EU27 & UK (*)" %in% unique(plotData_SEE_Total$region)) {
  plotData_SEE_Total <- plotData_SEE_Total %>% filter(region == "EU27 & UK (*)")
}
plotData_SEE_Total <- plotData_SEE_Total %>%
  mutate(model = as.character(model)) %>%
  select(model, period, value) %>%
  bind_rows(histStackedBars(data.frame(variable = "SE|Electricity", stringsAsFactors = FALSE),
                            histModel = "Ember")) %>%
  mutate(model = factor(model, levels = elecModelOrder))

g[["secondary_energy_electricity_stacked"]] <- ggplot(plotData_SEE) +
  geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
  geom_bar(aes(x = model, y = value, fill = category), stat = 'identity', alpha = 0.8) +
  geom_point(data = plotData_SEE_Total, aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
  theme_paper() +
  facet_wrap(~period, nrow = 1, scales = "free_x") +
  scale_x_discrete(limits = orderElecModels) +
  scale_fill_manual(values = elec_colors) +
  coord_cartesian(clip = 'off') +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
    strip.text = element_text(size = 12, colour = "black"),
    plot.margin = margin(t = 15, r = 25, b = 15, l = 20)
  ) +
  ylab("Secondary Energy | Electricity (EJ/yr)") +
  labs(title = "Electricity Generation by Source") +
  force_panelsizes(cols = panelBarCounts(plotData_SEE, plotData_SEE_Total))

ggsave(paste0(outFolder, "/svg/secondary_energy_electricity_stacked.svg"), g[["secondary_energy_electricity_stacked"]], device = "svg", width = 12, height = 8, dpi = 100, units = "in")
ggsave(paste0(outFolder, "/png/secondary_energy_electricity_stacked.png"), g[["secondary_energy_electricity_stacked"]], device = "png", width = 12, height = 8, dpi = 100, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["secondary_energy_electricity_stacked"]]

```

---

### Capacity | Electricity


Historical source: Ember, plus IRENA for geothermal
- PRIMES does not report electric capacity at all and LIMES only reports the total, so those two bars stay empty.
- Storage is left out of the stack: its power capacity is not part of the reported "Capacity|Electricity" total, so including it would push the bars above the diamonds.
- Ember covers every category except geothermal (about 0.9 GW in the EU), which comes from IRENA. The two combine cleanly - Ember's renewables capacity is IRENA data anyway, hydro, solar and wind match to four decimals.
- No source reports a usable historical capacity total across 2018-2022, so the historical bars carry no total diamond. Eurostat stops at 2018; the IEA WEO 2025 gives 2023 but is scope-inconsistent against Ember (solar matches to 0.02% while hydro is +67% and coal +52%), so it is not used.


```{r capacity_electricity_stacked, echo=FALSE, include=FALSE}

# Storage is left out of the stack: its power capacity is not part of the reported
# "Capacity|Electricity" total, so including it would push the bars above the diamonds.
# PRIMES does not report electric capacity at all and LIMES only reports the total, so
# those two bars stay empty.
cap_elec_vars <- setNames(paste0("Capacity|Electricity|", names(elec_colors)), names(elec_colors))

# Ember again, for consistency with the generation chart, plus IRENA for geothermal.
# No source reports a historical capacity total across 2018-2022 - Eurostat stops at 2018
# and the IEA WEO only gives 2023 - so the historical bars carry no total diamond.
cap_elec_hist_map <- bind_rows(
  data.frame(
    variable = paste0("Cap|Electricity|", c("Coal", "Oil", "Gas", "Biomass", "Nuclear", "Hydro", "Wind", "Solar")),
    category = c("Coal", "Oil", "Gas", "Biomass", "Nuclear", "Hydro", "Wind", "Solar"),
    source   = "Ember",
    stringsAsFactors = FALSE
  ),
  # Ember has no geothermal; IRENA does, and Ember's renewables capacity is IRENA data
  # anyway (hydro, solar and wind match to four decimals), so the two combine cleanly.
  # It is only ~0.9 GW of ~1030 GW, but it completes the category set.
  data.frame(variable = "Cap|Electricity|Geothermal", category = "Geothermal",
             source = "IRENA", stringsAsFactors = FALSE)
)

plotData_CAP <- df %>%
  filter(
    variable %in% cap_elec_vars,
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsElecall
  )

if ("EU27 & UK (*)" %in% unique(plotData_CAP$region)) {
  plotData_CAP <- plotData_CAP %>% filter(region == "EU27 & UK (*)")
}

capLabelMap <- setNames(names(cap_elec_vars), cap_elec_vars)
plotData_CAP <- plotData_CAP %>%
  mutate(model = as.character(model), category = capLabelMap[as.character(variable)]) %>%
  select(model, period, category, value) %>%
  bind_rows(histStackedBars(cap_elec_hist_map, histModel = "Ember")) %>%
  mutate(category = factor(category, levels = names(elec_colors)),
         model = factor(model, levels = elecModelOrder))

plotData_CAP_Total <- df %>%
  filter(
    variable == "Capacity|Electricity",
    period %in% c(2020, 2030, 2040, 2050),
    scenario == "WP1 NetZero",
    model %in% modelsElecall
  )
if ("EU27 & UK (*)" %in% unique(plotData_CAP_Total$region)) {
  plotData_CAP_Total <- plotData_CAP_Total %>% filter(region == "EU27 & UK (*)")
}
plotData_CAP_Total <- plotData_CAP_Total %>%
  mutate(model = factor(as.character(model), levels = elecModelOrder)) %>%
  select(model, period, value)

g[["capacity_electricity_stacked"]] <- ggplot(plotData_CAP) +
  geom_hline(yintercept = 0, color = "black", linewidth = 1, alpha = 0.5) +
  geom_bar(aes(x = model, y = value, fill = category), stat = 'identity', alpha = 0.8) +
  geom_point(data = plotData_CAP_Total, aes(x = model, y = value), colour = "black", size = 3, shape = 18) +
  theme_paper() +
  facet_wrap(~period, nrow = 1, scales = "free_x") +
  scale_x_discrete(limits = orderElecModels) +
  scale_fill_manual(values = elec_colors) +
  coord_cartesian(clip = 'off') +
  theme(
    axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = 0.5),
    strip.text = element_text(size = 12, colour = "black"),
    plot.margin = margin(t = 15, r = 25, b = 15, l = 20)
  ) +
  ylab("Capacity | Electricity (GW)") +
  labs(title = "Electric Capacity by Source") +
  force_panelsizes(cols = panelBarCounts(plotData_CAP, plotData_CAP_Total))

ggsave(paste0(outFolder, "/svg/capacity_electricity_stacked.svg"), g[["capacity_electricity_stacked"]], device = "svg", width = 12, height = 8, dpi = 100, units = "in")
ggsave(paste0(outFolder, "/png/capacity_electricity_stacked.png"), g[["capacity_electricity_stacked"]], device = "png", width = 12, height = 8, dpi = 100, units = "in")

```

```{r, echo=FALSE, fig.width=12, fig.height=8}

g[["capacity_electricity_stacked"]]

```







